### 2025-02-07 Rlearner_python_in_R.qmd
Python in R torch

---
title: "Rlearner_in_R"
format: html
editor: visual
echo: false
---

## Set up

```{r setup, include=FALSE}

# libraries
library(knitr)
library(readr)
library(torch)
library(reshape2)
library(dplyr)
library(GGally)
library(ggplot2)
library(purrr)

# directory 
knitr::opts_knit$set(root.dir = "/Users/monamousavi/Dropbox/Causal_climate/My_own_Shared/Takashi_code/Write_python_in_R/Data")

# no warnings in the output
options(warn=-1)
```

# Rann_regression_20230413 copy_d2

## Import data

```{r}
# Import data

# setwd("/Users/monamousavi/Dropbox/Causal_climate/My_own_Shared/Takashi_code/Write_python_in_R/Data")
data_2nd_stage <- read_csv("./DG2_ss.csv")
evall_N_seq <- read_csv("./DG2_eval_data.csv")

```

```{r}

plot_loss <- function(history) {
  loss <- as_array(history$loss)
  val_loss <- as_array(history$val_loss)

  # data frame
  loss_data <- data.frame(
    epoch = seq_along(loss),
    loss = loss,
    val_loss = val_loss
  )

  loss_data_long <- melt(loss_data, id.vars = "epoch")

  ggplot(loss_data_long, aes(x = epoch, y = value, color = variable)) +
    geom_line(size = 1) +
    labs(x = "Epoch", y = "MAE", color = "Loss Type") +
    theme_minimal() +
    theme(legend.position = "top") +
    geom_point()
}

```

# Multiple input DNN

```{r}

dataset <- data_2nd_stage %>%
  select(y_tilde, x1, x2, T_1_tilde, T_2_tilde, T_3_tilde)

# Split data into train (80%) and test (20%)

train_indices <- sample(1:nrow(dataset), size = 0.8 * nrow(dataset))  
train_dataset <- dataset[train_indices, ] 
test_dataset <- dataset[-train_indices, ]

# Convert data to torch tensors
train_tensor <- torch_tensor(as.matrix(train_dataset))
test_tensor <- torch_tensor(as.matrix(test_dataset))

```

```{r}

# Pair plot 
ggpairs(train_dataset, diag = list(continuous = "density")) +
  theme_minimal()

```

```{r}

# features
train_features_tensor <- train_tensor[, 2:6, drop = FALSE]
test_features_tensor <- test_tensor[, 2:6, drop = FALSE]

# labels
train_labels_tensor <- train_tensor[, 1, drop = FALSE]   # First column (y_tilde) 
test_labels_tensor <- test_tensor[, 1, drop = FALSE]   # First column (y_tilde)

# compute mean and Std (from train data) for standardization
train_means <- torch_mean(train_features_tensor, dim = 1)
train_sds <- torch_std(train_features_tensor, dim = 1, unbiased = TRUE)

# standardize features (use training stats for both train and test data)
scaled_train_features_tensor <- (train_features_tensor - train_means) / train_sds
scaled_test_features_tensor <- (test_features_tensor - train_means) / train_sds

```

```{r}

#  R torch neural network model
RModel <- nn_module(
  initialize = function() {

    self$dense1_x <- nn_linear(2, 64)
    self$dense2_x <- nn_linear(64, 64)
    self$dense3_x <- nn_linear(64, 1)

    self$dense1_T <- nn_linear(2, 64)
    self$dense2_T <- nn_linear(64, 64)
    self$dense3_T <- nn_linear(64, 1)

    self$dense1_T3 <- nn_linear(2, 64)
    self$dense2_T3 <- nn_linear(64, 64)
    self$dense3_T3 <- nn_linear(64, 1)
  },

  forward = function(input) {

    x <- input[, 1:2]   # First two columns
    T <- input[, 3:5]   # Last three columns

    T1 <- T[, 1, drop=FALSE]   # First column
    T2 <- T[, 2, drop=FALSE]   # Second column
    T3 <- T[, 3, drop=FALSE]   # Third column

    # Pass x to dense layers
    dense1 <- x %>%
      self$dense1_x() %>%
      nnf_relu() %>%
      self$dense2_x() %>%
      nnf_relu() %>%
      self$dense3_x()

    dense1 <- dense1 * T1

    dense2 <- x %>%
      self$dense1_T() %>%
      nnf_relu() %>%
      self$dense2_T() %>%
      nnf_relu() %>%
      self$dense3_T()

    dense2 <- dense2 * T2

    dense3 <- x %>%
      self$dense1_T3() %>%
      nnf_relu() %>%
      self$dense2_T3() %>%
      nnf_relu() %>%
      self$dense3_T3()

    dense3 <- dense3 * T3

    # Sum the outputs
    output <- dense1 + dense2 + dense3
    return(output)
  }
)

# model 
model <- RModel()

# optimizer and loss function
optimizer <- optim_adam(model$parameters, lr = 0.001)
loss_fn <- nnf_l1_loss  # MAE (Mean Absolute Error)

print(model)

```

```{r}

# Hyperparameters
epochs <- 500
batch_size <- 128
patience <- 10  # early stopping 
min_delta <- 0.0  # minimum improvement

# Optimizer & loss function
optimizer <- optim_adam(model$parameters, lr = 0.001)
loss_fn <- nnf_l1_loss  # MAE

train_dataset <- tensor_dataset(scaled_train_features_tensor, train_labels_tensor)
test_dataset <- tensor_dataset(scaled_test_features_tensor, test_labels_tensor)

train_dataloader <- dataloader(train_dataset, batch_size = batch_size, shuffle = TRUE)
test_dataloader <- dataloader(test_dataset, batch_size = batch_size, shuffle = FALSE)

# early stopping parameters
best_val_loss <- Inf
epochs_without_improvement <- 0

# history list
history <- list(train_loss = c(), val_loss = c())

# training Loop
for (epoch in 1:epochs) {
  model$train()
  train_loss <- 0

  coro::loop(for (batch in train_dataloader) {
    optimizer$zero_grad()

    inputs <- batch[[1]]
    targets <- batch[[2]]

    outputs <- model(inputs)
    loss <- loss_fn(outputs, targets)

    loss$backward()
    optimizer$step()

    train_loss <- train_loss + loss$item()
  })

  # average training loss
  train_loss <- train_loss / length(train_dataloader)

  # validation
  model$eval()
  val_loss <- 0

  coro::loop(for (batch in test_dataloader) {
    inputs <- batch[[1]]
    targets <- batch[[2]]

    outputs <- model(inputs)
    loss <- loss_fn(outputs, targets)

    val_loss <- val_loss + loss$item()
  })

  # average validation loss
  val_loss <- val_loss / length(test_dataloader)

  history$train_loss <- c(history$train_loss, train_loss)
  history$val_loss <- c(history$val_loss, val_loss)

  cat(sprintf("Epoch %d: Train Loss = %.4f | Val Loss = %.4f\n", epoch, train_loss, val_loss))

  if (val_loss < best_val_loss - min_delta) {
    best_val_loss <- val_loss
    epochs_without_improvement <- 0  # Reset counter
  } else {
    epochs_without_improvement <- epochs_without_improvement + 1
  }

  if (epochs_without_improvement >= patience) {
    cat("Early stopping triggered. Training stopped.\n")
    break
  }
}
```

```{r}

library(ggplot2)
library(tidyr)

plot_loss <- function(history) {
  # Convert history list to a data frame
  df <- data.frame(
    epoch = seq_along(history$train_loss),
    train_loss = history$train_loss,
    val_loss = history$val_loss
  ) %>%
    pivot_longer(cols = c(train_loss, val_loss), 
                 names_to = "Loss_Type", 
                 values_to = "Loss")

  ggplot(df, aes(x = epoch, y = Loss, color = Loss_Type)) +
    geom_line(linewidth = 1) +  # ✅ Fix applied
    geom_point() +
    labs(x = "Epoch", y = "Mean Absolute Error (MAE)", color = "Loss Type") +
    theme_minimal() +
    theme(legend.position = "top") +
    ggtitle("Training and Validation Loss")
}

plot_loss(history)

```

```{r}
# predictions (on test set)
model$eval()  
test_predictions_tensor <- model(scaled_test_features_tensor)  
test_predictions <- as_array(test_predictions_tensor)  # convert tensor to numeric vector
test_labels_vec <- as_array(test_labels_tensor)  # convert true labels to numeric vector

df <- data.frame(
  True_Values = test_labels_vec,
  Predictions = test_predictions
)

ggplot(df, aes(x = True_Values, y = Predictions)) +
  geom_point(color = "blue", alpha = 0.6) +  
  labs(x = "True Values", y = "Predictions", title = "model predictions vs true values") +
  theme_minimal() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")  
```

```{r}

# predict using the entire data

features <- dataset  
labels <- features$y_tilde  
features$y_tilde <- NULL

#  standardize features (use the same mean and std as training)
scaled_features <- scale(features, center = train_means, scale = train_sds)

# convert to torch tensor
scaled_features_tensor <- torch_tensor(as.matrix(scaled_features))
labels_tensor <- torch_tensor(as.numeric(labels))

#  Predictions
model$eval()  
predictions_tensor <- model(scaled_features_tensor)  
predictions <- as_array(predictions_tensor)  
true_values <- as_array(labels_tensor)

df <- data.frame(
  True_Values = true_values,
  Predictions = predictions
)

ggplot(df, aes(x = True_Values, y = Predictions)) +
  geom_point(color = "blue", alpha = 0.6) +  # Scatter plot
  labs(x = "True Values", y = "Predictions", title = "model predictions vs true values") +
  theme_minimal() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red")

```

```{r}
# loops over rows (repeats the row 100 times)

test_dataset <- data_2nd_stage
features <- test_dataset %>% select(x1, x2)
Tseq <- evall_N_seq$T

RMSE_vector <- c()

# iterate over 6 rows (like `for i in tqdm([0,1,2,3,4,5])` in takashi's)
walk(1:6, function(i) {
  feature_1 <- features[i, ]

  rep_feature_1 <- feature_1[rep(1, 100), ]  # repeat row
  rep_feature_1 <- cbind(rep_feature_1, evall_N_seq[, 1:3])  # T_1 and T_2 and T_3
  colnames(rep_feature_1) <- c("x1", "x2", "T_1_tilde", "T_2_tilde", "T_3_tilde")  # Rename columns

  # standardize features using the training mean and std
  scaled_feature_1 <- scale(rep_feature_1, center = train_means, scale = train_sds)
  scaled_feature_1_tensor <- torch_tensor(as.matrix(scaled_feature_1))  # Convert to tensor

  #  model predictions
  model$eval()  
  prediction_tensor <- model(scaled_feature_1_tensor)  
  prediction <- as_array(prediction_tensor)  # convert tensor to numeric vector

  # true_Y_tilde
  true_Y_tilde <- (feature_1$x1 + feature_1$x2^2) * log(Tseq)  # formula

  # predictions vs true values
  df <- data.frame(Tseq = Tseq, Prediction = prediction, True_Y = true_Y_tilde)

  p <- ggplot(df, aes(x = Tseq)) +
    geom_line(aes(y = Prediction, color = "Prediction")) +
    geom_line(aes(y = True_Y, color = "True_Y"), linetype = "dashed") +
    labs(x = "T", y = "Y_tilde", title = paste("Row", i, "Prediction vs True")) +
    theme_minimal() +
    scale_color_manual(values = c("Prediction" = "blue", "True_Y" = "black"))

   print(p)
  # compute RMSE
  MSE <- mean((true_Y_tilde - mean(true_Y_tilde) - (prediction - mean(prediction)))^2)
  RMSE <- sqrt(MSE)
  RMSE_vector <<- c(RMSE_vector, RMSE)  
})

print(RMSE_vector) # for 6 rows
```

# Rann_regression_MultiSimulation_train5field_20230504

## Import data

```{r}

data_2nd_stage <- readRDS("./data_20230504/data_2nd_stage.rds")
evall_N_seq = readRDS("./data_20230504/evall_N_seq.rds")
df_train_sim_id <- read_csv("./Output_ann_Train_5field/df_train_sim_id.csv")
```

```{r}

sim_id_train <- unlist(df_train_sim_id[2, ])   # row 1 in python is row 2 in R
sim_id_train <- sim_id_train[-1]

# convert to torch tensor and move to GPU (MPS)
sim_id_train_tensor <- torch_tensor(sim_id_train, device = torch_device("mps"))

# check it's on GPU
print(sim_id_train_tensor)

```

# check if GPU is working

```{r}
# check if GPU is working

mps_is_available <- function() {
  tryCatch({
    tmp <- torch_tensor(1, device = torch_device("mps"))
    TRUE
  }, error = function(e) {
    FALSE
  })
}

# --- Select device: CUDA > MPS > CPU ---
if (cuda_is_available()) {
  device <- torch_device("cuda")
} else if (mps_is_available()) {
  device <- torch_device("mps")
} else {
  device <- torch_device("cpu")
}
cat("Using device:", device$type, "\n")
```

# Multiple input DNN

```{r}

# loop for each simulation 
sim_ids <- unique(data_2nd_stage$sim)
num <- length(sim_ids)

for (j in seq_len(num)) {

  sim_ids_ <- sim_ids[-j]

  sim_id_train <- as.numeric(df_train_sim_id[2, ])
  sim_id_train <- sim_id_train[-1]  # remove first element
  train_sim_ids <- sim_id_train[1:2]

  sim_id_test <- sim_ids[j]
  cat("Processing sim_id_test =", sim_id_test, "\n")

  dataset <- data_2nd_stage %>%
    filter(sim %in% train_sim_ids) %>%
    select(y_tilde, Nk, plateau, b0, T_1_tilde, T_2_tilde, T_3_tilde)

  n <- nrow(dataset)
  train_idx <- sample(seq_len(n), size = floor(0.8 * n))
  train_dataset <- dataset[train_idx, ]
  val_dataset   <- dataset[-train_idx, ]

  train_labels <- train_dataset$y_tilde
  train_features <- train_dataset %>% select(-y_tilde)
  val_labels <- val_dataset$y_tilde
  val_features <- val_dataset %>% select(-y_tilde)

  # standardize

  col_means <- sapply(train_features, mean)
  col_sds   <- sapply(train_features, sd)

  scale_data <- function(df) {
    as.data.frame(scale(df, center = col_means, scale = col_sds))
  }

  scaled_train_features <- scale_data(train_features)
  scaled_val_features   <- scale_data(val_features)

  # convert data to torch tensors ---
  x_train <- torch_tensor(as.matrix(scaled_train_features), dtype = torch_float(), device = device)
  y_train <- torch_tensor(matrix(train_labels, ncol = 1), dtype = torch_float(), device = device)
  x_val   <- torch_tensor(as.matrix(scaled_val_features), dtype = torch_float(), device = device)
  y_val   <- torch_tensor(matrix(val_labels, ncol = 1), dtype = torch_float(), device = device)

  #  neural network model 
  Net <- nn_module(
    "Net",
    initialize = function() {
      # Branch 1
      self$branch1_fc1 <- nn_linear(3, 64)
      self$branch1_fc2 <- nn_linear(64, 64)
      self$branch1_fc3 <- nn_linear(64, 1)
      # Branch 2
      self$branch2_fc1 <- nn_linear(3, 64)
      self$branch2_fc2 <- nn_linear(64, 64)
      self$branch2_fc3 <- nn_linear(64, 1)
      # Branch 3
      self$branch3_fc1 <- nn_linear(3, 64)
      self$branch3_fc2 <- nn_linear(64, 64)
      self$branch3_fc3 <- nn_linear(64, 1)
    },
    forward = function(x) {
      x_main <- x[ , 1:3]
      T <- x[ , 4:6]

      # Branch 1
      b1 <- self$branch1_fc1(x_main) %>% nnf_relu()
      b1 <- self$branch1_fc2(b1) %>% nnf_relu()
      b1 <- self$branch1_fc3(b1)
      b1 <- b1 * T[ , 1, drop = FALSE]

      # Branch 2
      b2 <- self$branch2_fc1(x_main) %>% nnf_relu()
      b2 <- self$branch2_fc2(b2) %>% nnf_relu()
      b2 <- self$branch2_fc3(b2)
      b2 <- b2 * T[ , 2, drop = FALSE]

      # Branch 3
      b3 <- self$branch3_fc1(x_main) %>% nnf_relu()
      b3 <- self$branch3_fc2(b3) %>% nnf_relu()
      b3 <- self$branch3_fc3(b3)
      b3 <- b3 * T[ , 3, drop = FALSE]

      out <- b1 + b2 + b3
      out
    }
  )

  model <- Net()$to(device = device)

  optimizer <- optim_adam(model$parameters, lr = 0.001)

  # training loop 
  epochs <- 50     #500
  batch_size <- 512  
  patience <- 10
  best_val_loss <- Inf
  epochs_no_improve <- 0

  # when running the following code, I get this Warning:
  ### [W MPSFallback.mm:11] Warning: The operator 'aten::sgn.out' is not currently supported on the MPS backend and will fall back to run on the CPU. This may have performance implications. (function operator())

  for (epoch in 1:epochs) {
    model$train()
    optimizer$zero_grad()

    output <- model(x_train)
    loss <- nnf_l1_loss(output, y_train)

    loss$backward()
    optimizer$step()

    model$eval()
    with_no_grad({
      val_output <- model(x_val)
      val_loss <- nnf_l1_loss(val_output, y_val)$item()
    })

    # early stopping 
    if (val_loss < best_val_loss - 1e-6) {
      best_val_loss <- val_loss
      epochs_no_improve <- 0
    } else {
      epochs_no_improve <- epochs_no_improve + 1
    }

    if (epochs_no_improve >= patience) {
      cat("Early stopping at epoch", epoch, "\n")
      break
    }
  }

  ####$$%^^&&&***  End of where I get warning

  # predictions on val set 
  model$eval()
  with_no_grad({
    val_preds <- model(x_val)$to(device = "cpu")$numpy()
  })
  val_preds <- as.vector(val_preds)

  outcome_val <- data.frame(pred = val_preds, true = val_labels)
  outfile_val <- sprintf("./Output_Rann_Train_5field/validation_%s.csv", sim_id_test)
  write.csv(outcome_val, outfile_val, row.names = FALSE)

  # === EONR Estimation ===

  p_corn <- 6.25 / 25.4   
  p_N    <- 1 / 0.453592

  test_dataset <- data_2nd_stage %>% filter(sim == sim_id_test)

  test_features <- test_dataset %>% select(Nk, plateau, b0)

  test_evall_N_seq <- evall_N_seq %>%
    filter(sim == sim_id_test) %>%
    arrange(N)
  Nseq <- test_evall_N_seq$N
  n_row <- nrow(test_dataset)

  estEONR_vector <- numeric(n_row)

  # for each row (field) in the test data
  for (i in 1:n_row) {

    feature_1 <- test_features[i, , drop = FALSE]
    # repeat the row 100 times
    rep_feature_1 <- feature_1[rep(1, 100), ]

    temp <- test_evall_N_seq %>% select(T_1, T_2, T_3)
    names(temp) <- c("T_1_tilde", "T_2_tilde", "T_3_tilde")
    rep_feature_1 <- cbind(rep_feature_1, temp)

    rep_feature_1_scaled <- as.data.frame(mapply(function(col, mean_val, sd_val) {
      (col - mean_val) / sd_val
    }, rep_feature_1, col_means, col_sds, SIMPLIFY = FALSE))

    # convert the scaled data to a torch tensor.
    rep_feature_tensor <- torch_tensor(as.matrix(rep_feature_1_scaled), dtype = torch_float(), device = device)

    model$eval()
    with_no_grad({
      prediction <- model(rep_feature_tensor)$to(device = "cpu")$numpy()
    })
    prediction <- as.vector(prediction)

    MP <- prediction * p_corn - Nseq * p_N
    #  N value that maximizes MP
    estEONR <- Nseq[which.max(MP)]
    estEONR_vector[i] <- estEONR
  }

  outcome <- data.frame(pred = estEONR_vector, true = test_dataset$opt_N)
  outfile_eonr <- sprintf("./Output_Rann_Train_5field/EONR_%s.csv", sim_id_test)
  write.csv(outcome, outfile_eonr, row.names = FALSE)

}

```