### 2025-07-15 presentation/iisa-2025-TM.qmd

---
title: "Single-orthogonalized Neural Network for Site-specific Yield Response Function Estimation"
author:
  - name: Taro Mieno
    affiliations:
      - name: University of Nebraska Lincoln
  - name: Mona Mousavi
    affiliations:
      - name: University of Nebraska Lincoln
  - name: David Bullock
    affiliations:
      - name: University of Illinois Urbana Champaign 
date: 06/12/2025
format:
  beamer:
    theme: AnnArbor
format: 
  revealjs: 
    theme: [default, custom.scss]
    embed-resources: true
    html-math-method: mathjax
    fontsize: 1.4em
    callout-icon: false
    scrollable: true
    echo: true
---

```{r, include = FALSE}
library(terra)
library(sf)
library(tidyterra)
library(ggplot2)
library(data.table)
library(dplyr)
library(flextable)
```

## Background

+ Nitrogen rate management for agricultural production is an important issue (economic, environmental, health)

+ Applying right amount of nitrogen at right place (and right time) recognizing the need for nitrogen based on observed characteristics (soil N, soil and field characteristics, weather)

+ On-farm field experiment (a way)
  + run trials and collect data
  + <span style = "color: blue;"> estimate how yield responds to nitrogen based on observed characteristics </span>
  + identify site-specific optimal nitrogen rates based on the estimated model

+ Existing prediction oriented approaches (simple application of random forest, neural network) tend to fail for identifying the heterogeneous causal impact of a treatment

+ Existing causal machine learning approaches  (e.g., r-learner) specialize in identifying the impact of a binary treatment (but, we want yield response <span style = "color: blue;"> curve </span>)

```{=tex}
\begin{align*}
\Delta Yield = f(N_1| X) - f(N_0| X)\;\; \mbox{, where } N_1 > N_0
\end{align*}
```

## Research Objective

+ Propose an ML approach that can identify heterogenous treatment effects well when the treatment variable is continuous

+ Test the performance of the approach against existing approaches that are commonly used via Monte Carlo simulation

## Double machine learning and our approach

::: {.panel-tabset}

### Recap on Double-ML

```{=tex}
\begin{align*}
\mbox{Model 1: } Y = \theta(X)\cdot T + g(X, W) + v
\end{align*}
```

+ $Y$: Dependent variable 
+ $T$: Binary treatment variable (variable of interest) 
+ $X$: Set of variables that affect the impact of $T$ and also $Y$ directly
+ $\theta(X)$: Treatment effect as a function of $X$
+ $W$: Set of variables that affect $Y$ directly
+ $v$: Error term

::: {.fragment}
:::{.callout-note title="Steps"}
1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
2. Regress $T$ on $X$ and $W$ and predict $T[Y|X,W]$ (denoted as $h(X,W)$)
3. Regress $Y-f(X,W)$ on $T-h(X,W)$
:::
:::
<!--end of the fragment-->

### Single-orthogonalized (our approach)

```{=tex}
\begin{align*}
\mbox{Model 2: } Y = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ $T$: <span style = "color: blue;">Continuous </span> treatment variable (e.g., nitrogen rate) 
+ $\theta(X, T)$: Impact of $X$ and $T$ with their full interactions allowed

::: {.fragment}
:::{.callout-note title="Steps"}
1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
3. Regress $Y-f(X,W)$ on $T$
:::
:::
<!--end of the fragment-->

::: {.fragment}
:::{.callout-important title="Reasoning"}
+ The orthogonalization shifts the attention from the variations in $Y$ to the variations in $Y-f(X,W)$ just due to $X$ and $T$. 
+ This is particularly effective if the share of $g(X,W)$ in the total variation of $Y$ is greater relative to $\theta(X, T)$.
:::
:::
<!--end of the fragment-->

### Double-orthogonalized?

```{=tex}
\begin{align*}
\mbox{Model 1: } Y & = \theta(X)\cdot T + g(X, W) + v \\
\mbox{Model 2: } Y & = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ In model 1, Double-orthogonalization is fine because the model is linear in $T$.
+ In model 2, Orthogonalizing $T$ does not work: $\widehat{\theta}(X, T - g(X, W))$ cannot be used to approximate $E[\theta(X, T)]$.
:::
<!--end of panel-->

## Simulation: Data Generation

::: {.panel-tabset}

### Data generating process

Corn yield is assumed to follow the quadratic plateau functional form as follows:

```{=tex}
\begin{align*}
Y_{j, i}=f\left(N, \alpha_{j, i}, \alpha_{j, i}, \gamma{j, i}, \tau{j, i}\right)= \begin{cases}\alpha_{j, i}+\beta_{j, i} N+\gamma_{j, i} N^2+\varepsilon_{j, i}, & N<\tau_{j, i} \\ \alpha_{j, i}+\beta_{j, i} \tau_{j, i}+\gamma_{j, i} \tau_{j, i}^2+\varepsilon_{j, i}, & N \geq \tau_{j, i}\end{cases}
\end{align*}
```

+ $Y$: Corn yield
+ $N$: Nitrogen rate
+ $\alpha, \beta, \gamma, \tau$: parameters (consider them observable soil/field characteristics) that vary within a field in a spatially correlated manner and govern how yield responds to $N$

### On-farm trial (nitrogen rate)

```{r}
#| echo: false   
#| fig-width: 9
#| out-width: 9in
knitr::include_graphics("g_trial.png") 
```

### A parameter ($\tau$)

```{r}
#| echo: false   
#| fig-width: 9
#| out-width: 9in
knitr::include_graphics("g_par.png") 
```

### Corn yield

```{r}
#| echo: false   
#| fig-width: 9
#| out-width: 9in
knitr::include_graphics("g_yield.png") 
```

```{r}
#| echo: false
#| eval: false
data <- 
  readRDS(here::here("presentation/sim_data_1.rds"))$data[[1]] %>%
  sf::st_as_sf()

g_yield <- 
  ggplot(data = data) +
  geom_sf(aes(fill = yield)) +
  scale_fill_viridis_c(name = "Corn Yield (kg/ha)") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2.5, "cm"),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )

ggsave("presentation/g_yield.png", g_yield, dpi = 300, height = 4)

g_trial <-
  ggplot(data = data) +
  geom_sf(aes(fill = factor(N_tgt))) +
  scale_fill_viridis_d(name = "N rate (kg/ha)") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm"),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14)
  )

ggsave("presentation/g_trial.png", g_trial, dpi = 300, height = 4)

# library(latex2exp)
# plot(latex2exp::TeX("\\eta"))

g_par <-
  ggplot(data = data) +
  geom_sf(aes(fill = Nk)) +
  scale_fill_viridis_c(name = "") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(3, "cm"),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )

ggsave("presentation/g_par.png", g_par, dpi = 300, height = 4)
```

:::
<!--end of panel-->

## Simulation: Other Components

::: {.panel-tabset}

### Models

```{=tex}
\begin{align*}
Y_{i,k} = f(N_{i,k}, X_{i,k}) + v_{i,k}
\end{align*}
```

where $i$ and $k$ indicate a cell and field, respectively.

+ $Y$: Crop yield
+ $N$: Nitrogen rate (lb/acre) 
+ $X$: Collection of variables ($\alpha$, $\beta$, $\gamma$, and $\tau$)
+ $v$: Error term

<!-- ### Estimation approaches

<span style = "color: blue;"> Single-orthogonalized </span>

1. Regress $Y$ on $X$ ($N$ not included) using random forest, predict $Y$ conditional on $X$, and subtract it from $Y$ (orthogonalization)
2. Regress the orthogonalized $Y$ on $N$ and $X$ using neural network

<br>

<span style = "color: blue;">Non-orthogonalized  </span>

Regress $Y$ on $N$ and $X$ using neural network

<br>

<span style = "color: blue;">Double-orthogonalized </span>

1. Regress $Y$ on $X$ ($N$ not included) using random forest, predict $Y$ conditional on $X$, and subtract it from $Y$ (orthogonalization)
2. Regress $T$ on $X$ ($N$ not included) using random forest, predict $T$ conditional on $X$, and subtract it from $T$ (orthogonalization)
3. Regress the orthogonalized $Y$ on orthogonalized $N$ and $X$ using neural network -->

### Scenarios

+ Estimation approaches:

```{r}
#| echo: false
data.table(
  Approach = c("Double-orthogonalized", "Single-orthogonalized", "Non-orthogonalized"),
  "Y-orthogonalization" = c("Randome Forest", "Randome Forest", "NA"),
  "T-orthogonalization" = c("Randome Forest", "NA", "NA"),
  "Final stage" = c("Neural Network", "Neural Network", "Neural Network")
) %>%
flextable() %>%
fontsize(size = 16, part = "all") %>%
autofit()
```

+ Number of fields from which data is combined: 1, 5, and 10

### Site-specific EONR estimation

For a given price of corn and nitrogen (denoted $P_{Corn}$ and $P_{N}$, respectively), economically optimal nitrogen rate (EONR) is found for each cell of all the fields by solving the following profit maximization problem:

```{=tex}
\begin{align*}
\widehat{EONR}_{i,k} = Argmax_{N} \;\; P_{Corn} \times \widehat{f}(N, X_{i,k}) - P_{N} \times N
\end{align*}
```

+ $\widehat{f}$: estimated yield response curve

### Model evaluation

Model evaluation is based on the accuracy of predicting <span style = "color: blue;"> site-specific economically optimal nitrogen rates </span>($\widehat{EONR}$).

```{=tex}
\begin{align*}
RMSE = \sqrt{\sum_i^N \sum_k^K(EONR_{i,k} - \widehat{EONR_{i,k}})^2/NK}
\end{align*}
```

:::
<!--end of panel-->

## Results

## Conclusion