### 2025-07-15 presentation/iisa-2025-beamer-TM.qmd

---
title: "Single-orthogonalized Neural Network for Site-specific Yield Response Function Estimation"
author:
  - name: Taro Mieno$^*$
  - name: Mona Mousavi$^*$
  - name: David Bullock$^\dagger$
institute:
  - $^*$ University of Nebraska Lincoln
  - $^\dagger$ University of Illinois Urbana Champaign 
date: 06/12/2025
format:
  beamer:
    theme: Singapore
    navigation: horizontal
    section-titles: false
    slide-level: 2
    fontsize: "10pt"
header-includes: 
- \logo{\ifnum\thepage>1\includegraphics[width=1cm]{nebraska-n.jpg}\fi}
---

```{r, include = FALSE}
library(terra)
library(sf)
library(tidyterra)
library(ggplot2)
library(data.table)
library(dplyr)
library(flextable)
```

# Background and Research Objectives

## Background

+ Nitrogen rate management for agricultural production

+ Applying right amount of nitrogen at right place (and right time) recognizing the need for nitrogen based on observed characteristics (soil N, soil and field characteristics, weather)

+ On-farm field experiment (a way)
  + Run on-farm agronomic trials and collect data
  + \textcolor{blue}{Estimate how yield responds to nitrogen based on observed characteristics}
  + Identify site-specific optimal nitrogen rates

+ An increasing amount of data from on-farm field trials has been accumulated over the years.

+ Estimating yield response functions using data pooled from multiple fields can provide significant agronomic and economic insights.

## Motivation

+ When on-farm trial data is pooled, the treatment variable becomes "continuous" rather than discrete.

+ Existing approaches—such as direct applications of random forests or neural networks—often tend to fail to capture heterogeneous treatment effects of a continuous variable (Kakimoto et al., 2022).

+ Most causal machine learning methods (e.g., R-learner) are designed for binary treatments, whereas our objective is to recover the full yield response \textcolor{blue}{curve}.

## Research Objective

+ Find a causal machine learning approach capable of effectively identifying heterogeneous treatment effects tailored for yield response function estimation using data from on-farm trials, where the treatment variable is continuous and randomized.

+ Evaluate the performance of the proposed method through Monte Carlo simulations, comparing it against commonly used existing approaches.

# A New Approach

## Recap on Double-ML

```{=tex}
\begin{align*}
\mbox{Model 1: } Y = \theta(X)\cdot T + g(X, W) + v
\end{align*}
```

+ $Y$: Dependent variable 
+ $T$: Binary treatment variable (variable of interest) 
+ $X$: Set of variables that affect the impact of $T$ and also $Y$ directly
+ $\theta(X)$: Treatment effect as a function of $X$
+ $W$: Set of variables that affect $Y$ directly
+ $v$: Error term

. . .

:::{.callout-note title="Steps" icon=false}
1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
2. Regress $T$ on $X$ and $W$ and predict $T[Y|X,W]$ (denoted as $h(X,W)$)
3. Regress $Y-f(X,W)$ on $T-h(X,W)$
:::

## Single-orthogonalized (our approach)

```{=tex}
\begin{align*}
\mbox{Model 2: } Y = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ $T$: <span style = "color: blue;">Continuous </span> treatment variable (e.g., nitrogen rate) 
+ $\theta(X, T)$: Impact of $X$ and $T$ with their full interactions allowed

. . .

:::{.callout-note title="Steps" icon=false}
1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
3. Regress $Y-f(X,W)$ on $T$
:::

. . .

:::{.callout-important title="Reasoning" icon=false}
The orthogonalization shifts the attention from the variations in $Y$ to the variations in $Y-f(X,W)$ just due to $X$ and $T$. 
:::

## Double-orthogonalized?

```{=tex}
\begin{align*}
\mbox{Model 1: } Y & = \theta(X)\cdot T + g(X, W) + v \\
\mbox{Model 2: } Y & = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ In model 1, Double-orthogonalization is fine because the model is linear in $T$.
+ In model 2, Orthogonalizing $T$ does not work: $\widehat{\theta}(X, T - g(X, W))$ cannot be used to approximate $E[\theta(X, T)]$.

# Simulation

## Corn yield

Corn yield is assumed to follow the quadratic plateau functional form as follows:

```{=tex}
\begin{align*}
Y=f\left(N, \alpha, \beta{i, k}, \gamma, \tau\right)= \begin{cases}\alpha+\beta N+\gamma N^2+\varepsilon, & N<\tau \\ \alpha+\beta \tau+\gamma \tau^2+\varepsilon, & N \geq \tau\end{cases}
\end{align*}
```

+ $Y$: Corn yield
+ $N$: Nitrogen rate
+ $\alpha, \beta, \gamma, \tau$: parameters (consider them observable soil/field characteristics) that vary within a field in a spatially correlated manner and govern how yield responds to $N$

## On-farm trial (nitrogen rate)

```{r}
#| echo: false
#| fig-width: 4.5
#| out-width: 4.5in
knitr::include_graphics("g_trial.png")
```

## A parameter ($\tau$)

```{r}
#| echo: false
#| fig-width: 4.5
#| out-width: 4.5in
knitr::include_graphics("g_par.png")
```

## Corn yield

```{r}
#| echo: false
#| fig-width: 4.5
#| out-width: 4.5in
knitr::include_graphics("g_yield.png")
```

```{r}
#| echo: false
#| eval: false
data <-
  readRDS(here::here("presentation/sim_data_1.rds"))$data[[1]] %>%
  sf::st_as_sf()

g_yield <-
  ggplot(data = data) +
  geom_sf(aes(fill = yield)) +
  scale_fill_viridis_c(name = "Corn Yield (kg/ha)  ") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2, "cm"),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16)
  )

ggsave("presentation/g_yield.png", g_yield, dpi = 300, width = 10, height = 6)

g_trial <-
  ggplot(data = data) +
  geom_sf(aes(fill = factor(N_tgt))) +
  scale_fill_viridis_d(name = "N rate (kg/ha)") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18)
  )

ggsave("presentation/g_trial.png", g_trial, dpi = 300, width = 10, height = 6)

# library(latex2exp)
# plot(latex2exp::TeX("\\eta"))

g_par <-
  ggplot(data = data) +
  geom_sf(aes(fill = Nk)) +
  scale_fill_viridis_c(name = "") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(3, "cm"),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16)
  )

ggsave("presentation/g_par.png", g_par, dpi = 300, width = 10, height = 6)
```

## Models

```{=tex}
\begin{align*}
Y_{i,k} = f(N_{i,k}, X_{i,k}) + v_{i,k}
\end{align*}
```

where $i$ and $k$ indicate a cell and field, respectively.

+ $Y$: Crop yield
+ $N$: Nitrogen rate (lb/acre) 
+ $X$: Collection of variables ($\alpha$, $\beta$, $\gamma$, and $\tau$)
+ $v$: Error term

## Scenarios

+ Estimation approaches:

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("f_est.png")
```

```{r}
#| echo: false
#| eval: false
f_est <-
  data.table(
    Approach = c("Double-orthogonalized", "Single-orthogonalized", "Non-orthogonalized"),
    "Y-orthogonalization" = c("Randome Forest", "Randome Forest", "NA"),
    "T-orthogonalization" = c("Randome Forest", "NA", "NA"),
    "Final stage" = c("Neural Network", "Neural Network", "Neural Network")
  ) %>%
  flextable() %>%
  fontsize(size = 16, part = "all") %>%
  autofit()

flextable::save_as_image(f_est, here::here("presentation/f_est.png"), res = 300)
```

+ Number of fields from which data is combined: 1, 5, and 10

## Site-specific EONR estimation

For a given price of corn and nitrogen, economically optimal nitrogen rate (EONR) is found for each cell of all the fields by solving the following profit maximization problem:

```{=tex}
\begin{align*}
\widehat{EONR}_{i,k} = Argmax_{N} \;\; P_{Corn} \times \widehat{f}(N, X_{i,k}) - P_{N} \times N
\end{align*}
```

+ $P_{Corn}$: corn price
+ $P_N$: nitrogen price
+ $\widehat{f}$: yield response curve estimated by the model under evaluation

## Model evaluation

Model evaluation is based on the accuracy of predicting <span style = "color: blue;"> site-specific economically optimal nitrogen rates </span>($\widehat{EONR}$).

```{=tex}
\begin{align*}
RMSE = \sqrt{\sum_i^N \sum_k^K(EONR_{i,k} - \widehat{EONR_{i,k}})^2/NK}
\end{align*}
```

# Results and Conclusions

## Simulation Results

```{r}
#| eval: false
single_orth <-
  readRDS(here::here("results/Half_RANN_one_five_ten_fields.rds")) %>%
  tidyr::unnest(results) %>%
  data.table() %>%
  .[, est_type := "Single"]

double_orth <-
  readRDS(here::here("results/RANN_one_five_ten_fields.rds")) %>%
  tidyr::unnest(results) %>%
  data.table() %>%
  .[, est_type := "Double"]

no_orth <-
  readRDS(here::here("results/simple_ANN_one_five_ten_fields.rds")) %>%
  tidyr::unnest(results) %>%
  data.table() %>%
  .[, est_type := "Non"]

all_results <- rbind(single_orth, double_orth, no_orth)

rmse_data <-
  all_results[, .(rmse = sqrt(sum((true - pred)^2) / .N)), by = .(est_type, field, sim)] %>%
  .[, field_text := case_when(
    field == "1_field" ~ "Number of fields: 1",
    field == "5_fields" ~ "Number of fields: 5",
    field == "10_fields" ~ "Number of fields: 10",
  )] %>%
  .[, field_text := factor(field_text, levels = c("Number of fields: 1", "Number of fields: 5", "Number of fields: 10"))] %>%
  .[, est_type := factor(est_type, levels = c("Single", "Double", "Non"))]

g_results <-
  ggplot(data = rmse_data) +
  geom_histogram(aes(x = rmse), binwidth = 3) +
  facet_grid(est_type ~ field_text) +
  xlab("RMSE by iteration") +
  ylab("Count") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    strip.text = element_text(size = 20)
  )

ggsave("presentation/g_results.png", g_results, dpi = 300, height = 7.5)
```

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("g_results.png")
```

## Conclusion

+ The single-orthogonalized approach performed better than the non-orthogonalized (the common approach taken by researchers for yield response function estimation)
+ Double-orthogonalized approach performs poorly and its performance does not improve even if the number of observations increases (as expected)

## Future extensions

+ More realistic crop yield simulations using crop growth models
+ Other specifications of the approaches
  + Models other than NN for the final stage (e.g., random forest, boosted regression forest, etc)
  + Models other than RF for Y-orthogonalization
+ Is there any way to make the estimation more efficient taking into account the spatial correlation of the error term? (like spatial error model)