### 2025-07-15 codes/R/CF_Continuous_Loop.Rmd
minor edits

# Objectives

+ Check the CF performance by training on one, five and ten fields

# Preparation

## Packages and functions
```{r}
#--- packages ---#
library(tidyverse)
library(data.table)
library(mgcv)
library(spatialsample)
library(parallel)
library(sf)
library(grf)
library(ranger)
library(xgboost)
library(mgcv)
# library(here)
library(reticulate)

#--- R functions ---#
here("codes/R") %>%
  fs::dir_ls(., full.names = TRUE) %>%
  lapply(source)

#--- python functions ---#
use_condaenv("cf_conda311", required = TRUE)
py_config()
py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")
source_python("codes/Python/run_CF_c.py")
```

## Data

```{r}
# turn off multi-threading
Sys.setenv(OMP_NUM_THREADS = 1, MKL_NUM_THREADS = 1)

# load data only once, then drop
raw_sim_data <- readRDS("data/raw_sim_data.rds")
rm(list = setdiff(ls(), c("raw_sim_data", "np", "run_CF_c_py", "gm_wrapper_functions_you_need")))
gc()
```

# Train CF (Train on five fields. We can simply change 5 to other numbers)

```{r}
x_vars <- c("Nk", "plateau", "b0")
T_formula <- yield ~ s(N, k = 4)

output_dir <- "results/Seperate_Results/CF_in_R/all_sim_for_cf_5_fields"
output_dir <- "results/Seperate_Results/CF_in_R/all_sim_for_cf_5_fields"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

res_list_one <- lapply(1:500, function(test_sim_id) {
  result_file <- file.path(output_dir, paste0("sim_", test_sim_id, ".rds"))

  if (file.exists(result_file)) {
    message(sprintf("Skipping sim %d (already saved)", test_sim_id))
    return(NULL)
  }

  tryCatch(
    {
      # Train on five fields
      train_sim_ids <- sample(setdiff(1:500, test_sim_id), 5)
      sim_id_ls <- c(train_sim_ids, test_sim_id)

      data <- raw_sim_data$reg_data[[1]] %>%
        .[sim %in% sim_id_ls, ] %>%
        .[, .(sim, data)] %>%
        unnest(cols = c(data)) %>%
        data.table()

      train_data <- data[sim != test_sim_id, ]

      test_data <- raw_sim_data$reg_data[[1]] %>%
        .[sim %in% test_sim_id, ] %>%
        .[, .(sim, data)] %>%
        unnest(cols = c(data)) %>%
        data.table()

      T_info <- prepare_T_mat(T_formula, train_data)
      Y <- train_data[, yield]
      X <- train_data[, ..x_vars] %>% as.matrix()
      W <- X
      te_hat_cf <- run_CF_c_py(Y, T_info$T_sp, X, W, n_estimators = 2000)

      te_info <- get_te(te_hat_cf, test_data, x_vars)
      T_seq <- train_data[, seq(min(N), max(N), length.out = 200)]
      response_dt <- find_response_semi(T_seq, T_info, te_info)

      pCorn <- 6.25 / 25.4
      pN <- 1 / 0.453592

      ss_eonr <- response_dt[
        , profit := est * pCorn - pN * T
      ][
        , .SD[which.max(profit)],
        by = aunit_id
      ][
        , .(aunit_id, opt_N_hat = T)
      ]

      true_eonr <- raw_sim_data$reg_data[[1]] %>%
        .[sim == test_sim_id, ] %>%
        .[, .(sim, data)] %>%
        unnest(cols = c(data)) %>%
        data.table() %>%
        .[, .(aunit_id, b0, b1, b2, Nk)] %>%
        .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
        .[, opt_N := pmin(Nk, pmax(0, opt_N))] %>%
        .[, .(aunit_id, opt_N)]

      combined <- ss_eonr[true_eonr, on = "aunit_id"]
      combined[, sim := test_sim_id]

      saveRDS(combined, file = result_file)

      combined
    },
    error = function(e) {
      message(sprintf("Error with sim %d: %s", test_sim_id, e$message))
      return(data.table(sim = test_sim_id, aunit_id = integer(), opt_N_hat = numeric(), opt_N = numeric()))
    }
  )
})
```