### 2025-07-27 codes/R/Add_y_tilde_into_raw_sim_data.Rmd
Add Y_tilde into raw_sim_data

# add y_tilde to raw_sim_data

# load raw_sim_data and data_2nd_stage
```{r}
raw_sim_data <- readRDS("/Users/mmousavi2/Dropbox/Local-Cross-Validation/Shared/Data/raw_sim_data.rds")

data_2nd_stage <- readRDS("/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/data_20230504/data_2nd_stage.rds")

```

```{r}
# insert y_tilde into raw_sim_data
raw_sim_data$reg_data[[1]] <- raw_sim_data$reg_data[[1]] %>%
  mutate(
    data = pmap(
      list(data, sim),
      function(inner_df, sim_val) {
        inner_df %>%
          # temporarily add sim so we can join
          mutate(sim = sim_val) %>%
          left_join(
            data_2nd_stage %>% select(sim, aunit_id, y_tilde),
            by = c("sim", "aunit_id")
          ) %>%
          # remove the sim column—leave only the original columns + y_tilde
          select(-sim)
      }
    )
  )

```

# save it
```{r}
saveRDS(
  raw_sim_data,
  "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/raw_sim_data.rds"
)

```