### 2025-08-08 codes/R/Add_Ortho_N_to_data_2nd_stage.qmd
Add N_tilde to the data_2nd_stage

---
title: "Add N_tilde to the data_2nd_stage (<span style='color:blue'>Score: /10 </span>)"
author: "Your Name Here"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

# Load data
```{r setup_not_remove, echo=FALSE, include = F}

data_2nd_stage <- readRDS("/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/data_20230504/data_2nd_stage.rds")
evall_N_seq <- readRDS("/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/data_20230504/evall_N_seq.rds")

```

# Fit a RF
```{r setup_not_remove, echo=FALSE, include = F}

library(randomForest)

# fitting a RF to predict N from Nk, plateau, and b0

# rf_mod <- randomForest(
#   N ~ Nk + plateau + b0,
#   data = data_2nd_stage,
#   ntree = 100
# )

if (!requireNamespace("ranger", quietly = TRUE)) install.packages("ranger")
library(ranger)

rf_mod <- ranger(
  formula = N ~ Nk + plateau + b0,
  data    = data_2nd_stage,
  num.trees     = 100,       # or fewer if you like
  #mtry          = 2,         # smaller default than sqrt(p)
  min.node.size = 5,         # larger to speed up
  num.threads   = parallel::detectCores()
)

```

# add N_tilde to the data
```{r setup_not_remove, echo=FALSE, include = F}

# # Compute the predicted (fitted) values
# fitted_N <- predict(rf_mod, newdata = data_2nd_stage)
# 
# # Add the residualized treatment to the data 
# data_2nd_stage$N_tilde <- data_2nd_stage$N - fitted_N

data_2nd_stage$fitted_N <- predict(rf_mod, data = data_2nd_stage)$predictions
data_2nd_stage$N_tilde  <- data_2nd_stage$N - data_2nd_stage$fitted_N

```

# save data_2nd_stage with N_tilde
```{r setup_not_remove, echo=FALSE, include = F}
output_dir <- "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/data_20230504"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

saveRDS(data_2nd_stage,
        file = file.path(output_dir, "data_2nd_stage.rds"))

```

# add N_tilde seq to the eval_N_seq data
```{r setup_not_remove, echo=FALSE, include = F}
library(dplyr)

# Step 1: Create a sequence of 100 N_tilde values per sim
ntilde_seq <- data_2nd_stage %>%
  group_by(sim) %>%
  summarise(N_tilde = list(seq(min(N_tilde), max(N_tilde), length.out = 100))) %>%
  unnest(N_tilde)

# Step 2: Bind the new N_tilde column to evall_N_seq by sim
evall_N_seq <- evall_N_seq %>%
  arrange(sim) %>%
  group_by(sim) %>%
  mutate(row = row_number()) %>%
  ungroup()

ntilde_seq <- ntilde_seq %>%
  group_by(sim) %>%
  mutate(row = row_number()) %>%
  ungroup()

evall_N_seq <- evall_N_seq %>%
  left_join(ntilde_seq, by = c("sim", "row")) %>%
  select(-row)

```

# save eval_N_seq data with N_Tilde seq

```{r setup_not_remove, echo=FALSE, include = F}
output_dir <- "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/data/data_20230504"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

saveRDS(evall_N_seq,
        file = file.path(output_dir, "evall_N_seq.rds"))

```