### 2025-08-10 codes/R/debugging.Rmd
Check CF simulation to make sure there is no mistakes

# Objectives

+ Test whether python codes for estimating R-learner models are working properly using a dataset following a simple data generating process

+ Test whether spline approach (based on the `mgcv` package) is working properly

# Preparation

## Packages and functions
```{r}
#--- packages ---#
library(tidyverse)
library(data.table)
library(mgcv)
library(spatialsample)
library(parallel)
library(sf)
library(grf)
library(ranger)
library(xgboost)
library(mgcv)
#library(here)
#library(reticulate)
library(readr)
library(dplyr)
library(tidyr)

Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  VECLIB_MAXIMUM_THREADS = "1"  # macOS Accelerate safeguard
)

library(reticulate)
use_condaenv("cf_conda311", required = TRUE)
py_config()

py_run_string("
import platform, numpy as np, sklearn, scipy, econml
print('arch:', platform.machine())
print('numpy:', np.__version__)
print('sklearn:', sklearn.__version__)
print('scipy:', scipy.__version__)
print('econml:', econml.__version__)
")

# make sure Python can see the folder that contains run_CF_c.py
py_run_string("import sys, os; sys.path.append('/Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python')")

# clear any stale bytecode that might have been written under a different toolchain
system("find /Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python -name '__pycache__' -type d -prune -exec rm -rf {} +")

# now import without executing R helpers:
py_run_string("import run_CF_c; print('run_CF_c imported OK; has run_CF_c_py:', hasattr(run_CF_c, 'run_CF_c_py'))")

#--- python functions ---#
# use_condaenv("cf_conda311", required = TRUE)
# py_config()
# py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")
source_python("/Users/mmousavi2/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python/run_CF_c.py")
```

## Data

```{r}
all_data <- raw_sim_data

test_sim_id=1

split_row     <- train_test_splits_10fields[train_test_splits_10fields$test_id == test_sim_id, ]
train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:10)]))

sim_id_ls     <- c(train_sim_ids, test_sim_id)

    # assemble train & test data
    data <- raw_sim_data$reg_data[[1]] %>%
      .[sim %in% sim_id_ls, ] %>%
      .[, .(sim, data)] %>%
      unnest(cols = c(data)) %>%
      data.table()

    train_data <- data[sim != test_sim_id, ]

    # aunit level
    test_data  <- raw_sim_data$reg_data[[1]] %>%
      .[sim == test_sim_id, ] %>%
      .[, .(sim, data)] %>%
      unnest(cols = c(data)) %>%
      data.table()

    # cell level
  # test_data <-
  # all_data$field_pars[[1]] %>%
  # .[sim == test_sim_id, ]

```

# Train CF

```{r, eval = FALSE}
x_vars <- c("Nk", "plateau", "b0")
# x_vars <-
#   c(
#     "b0_1", "b0_2",
#     "b1_2_1", "b1_2_2", "b1_1_1", "b1_1_2",
#     # "b2_2_1", "b2_2_2", "b2_1_1", "b2_1_2",
#     "Nk_2_1", "Nk_2_2", "Nk_1_1", "Nk_1_2",
#     "plateau_2_1", "plateau_2_2", "plateau_1_1", "plateau_1_2"
#     # "theta_plateau_2", "theta_Nk_2", "theta_b0_2"
#   )

T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
Y <- train_data[, yield]
X <- train_data[, ..x_vars] %>% as.matrix()
W <- X
X_test <- test_data[, ..x_vars] %>% as.matrix()

te_hat_cf <- run_CF_c_py(Y, T_info$T_sp, X, W, n_estimators = 2000)

te_info <- get_te(te_hat_cf, test_data, x_vars, "aunit_id")
#te_info <- get_te(te_hat_cf, test_data, x_vars, "cell_id")

T_seq <- train_data[, seq(min(N), max(N), length = 200)]

response_data <- find_response_semi(T_seq, T_info, te_info)

saveRDS(list(te_hat_cf = te_hat_cf, te_info = te_info, T_seq = T_seq, response_data = response_data, test_data = test_data, train_data = train_data), "/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/results/Test_CF_any_mis/cf-continuous-unit_10_fields_double_check.rds")
```

# Post-estimation analysis
```{r}
cf_results <- readRDS("/Users/mmousavi2/Dropbox/Causal_climate/Data_and_results_CF_Continuous/results/Test_CF_any_mis/cf-continuous-unit_10_fields_double_check.rds")
```

## Site-specific EONR

```{r}
pCorn <- 6.25 / 25.4 # $/kg
pN <- 1 / 0.453592 # $/kg

# aunit id ###############################
ss_eonr <-
  cf_results$response_data %>%
  .[, profit := est * pCorn - pN * T] %>%
  .[, .SD[which.max(profit), ], by = aunit_id] %>%
  .[, .(aunit_id, T)] %>%
  setnames("T", "opt_N_hat")

true_ss_eonr <-
  cf_results$test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
  .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
  .[, opt_N := pmin(Nk, opt_N)] %>%
  .[, opt_N := pmax(0, opt_N)] %>%
  .[, .(aunit_id, opt_N)]

combined_eonr <- ss_eonr[true_ss_eonr, on = "aunit_id"]

combined_eonr[, cor(opt_N, opt_N_hat)]

############################################################################

# cell ###########################
# ss_eonr <-
#   cf_results$response_data %>%
#   .[, profit := est * pCorn - pN * T] %>%
#   .[, .SD[which.max(profit), ], by = cell_id] %>%
#   .[, .(cell_id, T)] %>%
#   setnames("T", "opt_N_hat")
#  
#  true_ss_eonr <-
#   cf_results$test_data[, .(cell_id, b0, b1, b2, Nk)] %>%
#   .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
#   .[, opt_N := pmin(Nk, opt_N)] %>%
#   .[, opt_N := pmax(0, opt_N)] %>%
#   .[, .(cell_id, opt_N)]
#  
# combined_eonr <- ss_eonr[true_ss_eonr, on = "cell_id"]

#combined_eonr[, cor(opt_N, opt_N_hat)]

#######################################################################################

rmse_10_fields_unit <- sqrt(mean((combined_eonr$opt_N_hat - combined_eonr$opt_N)^2, na.rm = TRUE))  
rmse_10_fields_unit
```

# dependency functions
```{r}
prepare_T_mat <- function(gam_formula, data) {
  gam_setup <- gam(gam_formula, data = data)
  T_var_name <- all.vars(gam_formula)[2]
  T_sp <- predict(gam_setup, data = data, type = "lpmatrix") %>%
    .[, -1] %>%
    data.table() %>%
    setnames(names(.), paste0("T_", 1:ncol(.)))
  return(list(gam_setup = gam_setup, T_sp = T_sp, T_var_name = T_var_name))
}

################################################################################
# aunit

get_te <- function(trained_model, test_data, x_vars, id_var = "aunit_id") {
  X_test <- test_data[, ..x_vars] %>% as.matrix()
  te_hat <- trained_model$const_marginal_effect(X_test)
  return(list(te_hat = data.table(te_hat), id_data = test_data[, ..id_var]))
}

# cell

# get_te <- function(trained_model, test_data, x_vars, id_var = "cell_id") {
#   X_test <- test_data[, ..x_vars] %>% as.matrix()
#   te_hat <- trained_model$const_marginal_effect(X_test)
#   return(list(te_hat = data.table(te_hat), id_data = test_data[, ..id_var]))
# }
################################################################################
find_response_semi <- function(T_seq, T_info, te_info) {
  eval_T <- data.table(T = T_seq) %>%
    setnames("T", T_info$T_var_name) %>%
    predict(T_info$gam_setup, newdata = ., type = "lpmatrix") %>%
    .[, -1] %>%
    data.table()

  curv_data <- as.matrix(te_info$te_hat) %*% t(as.matrix(eval_T)) %>%
    data.table() %>%
    setnames(names(.), as.character(T_seq)) %>%
    .[, id := 1:.N] %>%
    melt(id.var = "id") %>%
    setnames(c("variable", "value"), c("T", "est")) %>%
    .[, T := as.numeric(as.character(T))]

  id_data <- te_info$id_data[, id := 1:.N]
  final_data <- id_data[curv_data, on = "id"][, id := NULL]
  return(final_data)
}

```