### 2025-08-14 codes/R/CF_loop.Rmd

## Packages and functions
```{r}
#--- packages ---#
library(tidyverse)
library(data.table)
library(mgcv)
library(spatialsample)
library(parallel)
library(sf)
library(grf)
library(ranger)
library(xgboost)
library(mgcv)
# library(here)
library(reticulate)
library(readr)
library(dplyr)
library(tidyr)

#--- R functions ---#
# here("codes/R") %>%
#   fs::dir_ls(., full.names = TRUE) %>%
#   lapply(source)

use_virtualenv("cf_env311", required = TRUE)
Sys.setenv(DYLD_LIBRARY_PATH = "/opt/homebrew/Cellar/llvm@15/15.0.7/lib/c++")
py_config()

source_python("/Users/monamousavi/Dropbox/Causal_climate/My_own_Shared/Causal-ML-continuous-treatment/codes/Python/run_CF_c.py")
```

# CF loop

```{r}
## ───────────────────────────────────────
train_test_splits <- train_test_splits_1fields
n_fields <- 1 # how many train sims

## ── Output paths ──────────────────────────
output_dir <- "/Users/monamousavi/Dropbox/CF_res_1"
dir.create(path.expand(output_dir), recursive = TRUE, showWarnings = FALSE)
progress_csv <- file.path(output_dir, "progress.csv")

## ─  keep an in-memory list ────────────
# cf_all_results <- vector("list", length = nrow(train_test_splits))
# names(cf_all_results) <- train_test_splits$test_id
train_test_splits <- subset(train_test_splits, !test_id %in% as.integer(sub("^sim_(\\d+)\\.rds$", "\\1", basename(list.files(path.expand(output_dir), "^sim_\\d+\\.rds$")))))

## ── Loop over test sims ───────────────────────────────────────────────────────
for (i in seq_len(nrow(train_test_splits))) {
  test_sim_id <- train_test_splits$test_id[i]
  out_fn <- file.path(output_dir, sprintf("sim_%03d.rds", as.integer(test_sim_id)))

  # skip if already saved
  if (file.exists(out_fn)) {
    message("Skipping sim ", test_sim_id, " (already saved)")
    next
  }

  #
  tryCatch(
    {
      ## 1) Train/test IDs
      split_row <- train_test_splits[i, ]
      train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:n_fields)]))
      sim_id_ls <- c(train_sim_ids, test_sim_id)

      ## 2) train & test data
      data <- raw_sim_data$reg_data[[1]] %>%
        .[sim %in% sim_id_ls, ] %>%
        .[, .(sim, data)] %>%
        tidyr::unnest(cols = c(data)) %>%
        as.data.table()

      train_data <- data[sim != test_sim_id, ]
      test_data <- raw_sim_data$reg_data[[1]] %>%
        .[sim == test_sim_id, ] %>%
        .[, .(sim, data)] %>%
        tidyr::unnest(cols = c(data)) %>%
        as.data.table()

      ## 3)  CF
      x_vars <- c("Nk", "plateau", "b0")
      T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
      Y <- train_data[, yield]
      X <- as.matrix(train_data[, ..x_vars])
      W <- X
      X_test <- as.matrix(test_data[, ..x_vars])

      ## 4) Run CF in Python
      Sys.setenv(
        OMP_NUM_THREADS        = "1",
        MKL_NUM_THREADS        = "1",
        KMP_AFFINITY           = "disabled",
        KMP_DUPLICATE_LIB_OK   = "TRUE"
      )

      te_hat_cf <- run_CF_c_py(
        Y,
        T_info$T_sp,
        X,
        W,
        n_estimators = 2000
      )
      te_info <- get_te(te_hat_cf, test_data, x_vars, "aunit_id")

      ## 5) Response curves
      T_seq <- train_data[, seq(min(N), max(N), length = 200)]
      response_data <- find_response_semi(T_seq, T_info, te_info)

      ## 6) Site-specific EONR
      pCorn <- 6.25 / 25.4
      pN <- 1 / 0.453592

      ss_eonr <- response_data %>%
        .[, profit := est * pCorn - pN * T] %>%
        .[, .SD[which.max(profit), ], by = aunit_id] %>%
        .[, .(aunit_id, opt_N_hat = T)]

      true_ss_eonr <- test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
        .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
        .[, opt_N := pmin(Nk, opt_N)] %>%
        .[, opt_N := pmax(0, opt_N)] %>%
        .[, .(aunit_id, opt_N)]

      combined_eonr <- ss_eonr[true_ss_eonr, on = "aunit_id"]
      corr_val <- combined_eonr[, cor(opt_N, opt_N_hat)]

      ## 7) Package result for THIS sim
      res <- list(
        te_hat_cf     = te_hat_cf,
        te_info       = te_info,
        T_seq         = T_seq,
        response_data = response_data,
        test_data     = test_data,
        train_data    = train_data,
        combined_eonr = combined_eonr,
        corr          = corr_val
      )

      #  keep in memory too
      # cf_all_results[[as.character(test_sim_id)]] <- res

      ## 8) SAVE THIS SIM IMMEDIATELY
      tmp_fn <- paste0(out_fn, ".tmp")
      saveRDS(res, tmp_fn, compress = "xz")
      file.rename(tmp_fn, out_fn)

      ## 9) Log progress
      prog <- data.frame(
        test_sim_id = test_sim_id,
        corr        = corr_val,
        n_train     = nrow(train_data),
        saved_file  = out_fn,
        ts          = as.character(Sys.time())
      )
      write.table(
        prog,
        file      = progress_csv,
        sep       = ",",
        row.names = FALSE,
        col.names = !file.exists(progress_csv),
        append    = TRUE
      )

      message(
        "Finished sim ", test_sim_id, " → ", basename(out_fn),
        " (corr = ", round(corr_val, 3), ")"
      )

      # Free RAM periodically
      rm(res, te_hat_cf, te_info, response_data, train_data, test_data)
      if (i %% 10 == 0) gc()
    },
    error = function(e) {
      message("ERROR in sim ", test_sim_id, ": ", conditionMessage(e))
      #  which sim failed
      err_row <- data.frame(
        test_sim_id = test_sim_id,
        error       = conditionMessage(e),
        ts          = as.character(Sys.time())
      )
      write.table(
        err_row,
        file      = file.path(output_dir, "errors.csv"),
        sep       = ",",
        row.names = FALSE,
        col.names = !file.exists(file.path(output_dir, "errors.csv")),
        append    = TRUE
      )
    }
  )
}

## ── save the whole list at the end ─────────────────────────────────
# saveRDS(cf_all_results, file.path(output_dir, "_all_results.rds"))
```

# read in

```{r}
library(data.table)

in_dir <- "/Users/monamousavi/Dropbox/CF_res_1" # where current sim_###.rds live
out_dir <- "~/Desktop/CF_res_1_eonr" # new lean outputs
dir.create(path.expand(out_dir), recursive = TRUE, showWarnings = FALSE)

files <- list.files(path.expand(in_dir), pattern = "^sim_\\d+\\.rds$", full.names = TRUE)
sim_id <- as.integer(sub("^.*sim_(\\d+)\\.rds$", "\\1", files))

for (i in seq_along(files)) {
  r <- readRDS(files[i]) # don't View(); just extract the piece we need
  e <- as.data.table(r$combined_eonr)[, .(aunit_id, opt_N_hat, opt_N)]
  e[, sim := sim_id[i]]
  setcolorder(e, c("aunit_id", "opt_N_hat", "opt_N", "sim"))

  fn_rds <- file.path(out_dir, sprintf("sim_%03d_eonr.rds", sim_id[i]))
  tmp <- paste0(fn_rds, ".tmp")
  saveRDS(e, tmp, compress = "xz")
  file.rename(tmp, fn_rds)

  # optional CSV
  # fwrite(e, sub("\\.rds$", ".csv", fn_rds))

  rm(r, e)
  if (i %% 50 == 0) gc()
}
```