### 2025-08-17 codes/R/YRF_visualized.qmd
present results (yield response)

---
title: "yield response function"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

## Packages and functions
```{r setup_not_remove, echo=FALSE, include = F}
#--- R functions ---#
library(here)
source(here("codes", "R", "visualize_yield_response_function.R"))
```

# Use it:

```{r}
plot_yield_response(model = "NO_ANN", n_fields = 5)     # uses N
plot_yield_response(model = "DO_ANN", n_fields = 5)     # uses N_tilde 
plot_yield_response(model = "SO_ANN", n_fields = 10) 
plot_yield_response(model = "NO_RF", n_fields = 5) 
plot_yield_response(model = "SO_RF", n_fields = 5) 
```