### 2025-08-19 codes/R/visualized_results.qmd
minor edits

---
title: "yield response function"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

## Packages and functions
```{r setup_not_remove, echo=FALSE, include = F}
#--- R functions ---#
library(here)
source(here("codes", "R", "visualize_yield_response_function.R"))
```

# Use it:

```{r}
plot_yield_response(model = "NO_ANN", n_fields = 5)     # uses N
plot_yield_response(model = "DO_ANN", n_fields = 5)     # uses N_tilde 
plot_yield_response(model = "SO_ANN", n_fields = 10) 
plot_yield_response(model = "NO_RF", n_fields = 5) 
plot_yield_response(model = "SO_RF", n_fields = 5) 
```
# RMSE

```{r}
# ============================
# create per-sim RMSE table 
# ============================

# ---- Packages ----
library(tidyverse)
library(fs)
library(here)
library(readr)
library(stringr)

# ---- Constants: all 18 result folders ----
MODEL_DIRS <- c(
  "CF_outcome_1_fields","CF_outcome_5_fields","CF_outcome_10_fields",
  "RANN_outcome_one_field","RANN_outcome_five_fields","RANN_outcome_ten_fields",
  "Simple_ANN_outcome_one_field","Simple_ANN_outcome_five_fields","Simple_ANN_outcome_ten_fields",
  "NO_RF_one_field","NO_RF_five_fields","NO_RF_ten_fields",
  "HalfRRF_one_field","HalfRRF_five_fields","HalfRRF_ten_fields",
  "half_RANN_outcome_one_field","half_RANN_outcome_five_fields","half_RANN_outcome_ten_fields"
)

EXPECTED_ROWS <- 1440L

# ---- function ----
rmse <- function(x, y) sqrt(mean((x - y)^2, na.rm = TRUE))

parse_n_fields <- function(folder) {
  case_when(
    str_detect(folder, "(?:_1_fields|one_field)\\b")   ~ 1L,
    str_detect(folder, "(?:_5_fields|five_fields)\\b") ~ 5L,
    str_detect(folder, "(?:_10_fields|ten_fields)\\b") ~ 10L,
    TRUE ~ NA_integer_
  )
}

parse_model <- function(folder) {
  f <- str_to_lower(folder)
  case_when(
    str_starts(f, "cf_")             ~ "CF",
    str_starts(f, "rann_outcome")    ~ "RANN",
    str_starts(f, "half_rann")       ~ "HalfRANN",
    str_starts(f, "simple_ann")      ~ "Simple_ANN",
    str_starts(f, "no_rf")           ~ "NO_RF",
    str_starts(f, "halfrrf")         ~ "HalfRRF",
    TRUE                             ~ folder
  )
}

# Read any object: RDS, CSV
read_any <- function(f) {
  ext <- tolower(path_ext(f))
  if (ext == "csv") return(read_csv(f, show_col_types = FALSE))
  obj <- tryCatch(readRDS(f), error = function(e) NULL)
  if (!is.null(obj)) return(obj)
  env <- new.env(parent = emptyenv())
  ok <- tryCatch({ nm <- load(f, envir = env); length(nm) > 0 }, error = function(e) FALSE)
  if (ok) return(env[[ls(env)[1]]])
  stop("Unsupported or unreadable file: ", f)
}

# Standardize to two numeric vectors (per-sim rows)
read_eonr_file <- function(f) {
  obj <- read_any(f)
  if (is.list(obj) && all(c("pred","true") %in% names(obj))) {
    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
  } else if (is.data.frame(obj) && all(c("pred","true") %in% names(obj))) {
    list(pred = as.numeric(obj$pred), true = as.numeric(obj$true))
  } else if (is.data.frame(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF
  } else if (is.list(obj) && all(c("opt_N_hat","opt_N") %in% names(obj))) {
    list(pred = as.numeric(obj$opt_N_hat), true = as.numeric(obj$opt_N))  # CF list
  } else {
    stop("Unknown structure (need pred/true or opt_N_hat/opt_N): ", f)
  }
}

# 
files_in_dir <- function(dir_path) {
  dir_ls(dir_path, type = "file", recurse = TRUE) %>%
    # Names like sim_055_eonr or EONR_1 
    keep(~ str_detect(path_file(.x),
           regex("^(sim_\\d+(?:_eonr)?|EONR_\\d+)(?:\\.(rds|rda|rdata|csv))?$",
                 ignore_case = TRUE))) %>%
    # drop anything starting with "validation"
    discard(~ str_detect(path_file(.x), regex("^validation", ignore_case = TRUE))) %>%
    # drop macOS Icon files
    discard(~ grepl("Icon", basename(.x), fixed = TRUE))
}

# get sim id from filename base
extract_sim_id <- function(base) {
  m <- str_match(base, regex("^(?:sim_(\\d+)(?:_eonr)?|EONR_(\\d+))", ignore_case = TRUE))
  coalesce(as.integer(m[,2]), as.integer(m[,3]))
}

# 
row_from_file <- function(f, model, n_f, expected_rows = EXPECTED_ROWS) {
  tryCatch({
    base <- path_ext_remove(path_file(f))
    sim_id <- extract_sim_id(base)
    if (is.na(sim_id)) return(NULL)

    e <- read_eonr_file(f)
    n_rows <- length(e$pred)

    tibble(
      sim      = sim_id,
      Model    = model,
      `#field` = n_f,
      n_rows   = n_rows,
      RMSE     = rmse(e$pred, e$true),
      ok_rows  = isTRUE(n_rows == expected_rows),
      file     = f
    )
  }, error = function(e) NULL)
}

# ---- create table  ----
build_rmse_table <- function(base_dir = here("results"),
                             include_dirs = MODEL_DIRS,
                             expected_rows = EXPECTED_ROWS) {
  dirs <- file.path(base_dir, include_dirs)
  dirs <- dirs[dir_exists(dirs)]

  map_dfr(dirs, function(dir_path) {
    folder <- path_file(dir_path)
    model  <- parse_model(folder)
    n_f    <- parse_n_fields(folder)

    files <- files_in_dir(dir_path)

    map_dfr(files, row_from_file, model = model, n_f = n_f,
            expected_rows = expected_rows)
  }) %>%
    filter(!is.na(sim), !is.na(`#field`), !is.na(RMSE)) %>%
    arrange(Model, `#field`, sim)
}

```

# RMSE run

```{r}
# ---- run it (all models) ----
rmse_tbl_raw <- build_rmse_table(
  base_dir = here("results"),
  include_dirs = MODEL_DIRS,
  expected_rows = EXPECTED_ROWS
)

# save
# write_csv(rmse_tbl_raw, here("results", "rmse_table_all_models.csv"))
#saveRDS(rmse_tbl_raw, here("results", "rmse_tbl_raw.rds"))

```

# his

```{r}

library(dplyr)
library(ggplot2)
library(stringr)

#  model names: row labels in the graph
model_to_row <- c(
  "Simple_ANN" = "No_ANN",
  "RANN"       = "DO_ANN",
  "NO_RF"      = "NO_RF",
  "HalfRANN" = "SO_ANN",
  "HalfRRF"  = "SO_RF",
   "CF"       = "CF"
)

rmse_plot_df <- rmse_tbl_raw %>%
  filter(ok_rows, Model %in% names(model_to_row)) %>%
  mutate(
    row_lab = recode(Model, !!!model_to_row),
    col_lab = factor(`#field`,
                     levels = c(1, 5, 10),
                     labels = c("Number of fields: 1",
                                "Number of fields: 5",
                                "Number of fields: 10"))
  )

#  check 
rmse_plot_df %>% count(row_lab, col_lab)

# --- setup ---
xmin <- 0
nice_ceiling <- function(x, base = 5) base * ceiling(x / base)
xmax <- nice_ceiling(max(rmse_plot_df$RMSE, na.rm = TRUE), base = 5)

bins <- 25
binwidth <- (xmax - xmin) / bins

# --- Plot ---
p <- ggplot(rmse_plot_df, aes(x = RMSE)) +
  geom_histogram(binwidth = binwidth, boundary = 0, closed = "left") +
  facet_grid(row_lab ~ col_lab, scales = "fixed") +  
  scale_x_continuous(limits = c(xmin, xmax),
                     breaks = seq(xmin, xmax, length.out = 6),
                     expand = c(0, 0)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(x = "RMSE by iteration", y = "Count") +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold", size = 16),
    axis.title = element_text(size = 14)
  )

p

# ggsave(here::here("results", "rmse_hist_3rows.png"), p, width = 12, height = 7, dpi = 300)

```

# create rmse_summary

```{r}
# =========
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)

# 
rmse_df <- rmse_tbl_raw %>%
  dplyr::group_by(Model, `#field`, sim) %>%  
  dplyr::summarise(RMSE = dplyr::first(RMSE), .groups = "drop") %>%
  # Rename models
  dplyr::mutate(
    Model = dplyr::recode(
      Model,
      "Simple_ANN" = "NO_ANN",
      "HalfRANN"   = "SO_ANN",
      "RANN"       = "DO_ANN",
      "HalfRRF"    = "SO_RF",
      "CF"         = "CF",
      "NO_RF"      = "NO_RF",
      .default     = Model
    ),
    field_f = factor(`#field`, levels = c(1, 5, 10),
                     labels = c("1 field", "5 fields", "10 fields"))
  )

# 3) summary stats 
rmse_summary <- rmse_df %>%
  dplyr::group_by(Model, field_f) %>%
  dplyr::summarise(
    n      = dplyr::n(),
    mean   = mean(RMSE, na.rm = TRUE),
    median = median(RMSE, na.rm = TRUE),
    sd     = sd(RMSE, na.rm = TRUE),
    se     = sd / sqrt(n),
    lower  = mean + qt(0.025, df = n - 1) * se,
    upper  = mean + qt(0.975, df = n - 1) * se,
    .groups = "drop"
  )

```

# plot 1

```{r}
# 
p_mean <- ggplot(rmse_summary, aes(x = field_f, y = mean, group = Model, color = Model)) +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(size = 3) +
  labs(x = "# fields", y = "Mean RMSE", color = "Model") +
  theme_bw() +
  theme(
    axis.title = element_text(size = 13),
    strip.text = element_text(face = "bold")
  )

p_mean +
  geom_text(aes(label = sprintf("%.1f", mean)), vjust = -0.6, size = 3, show.legend = FALSE) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.10)))

p_mean

# ggsave(here::here("results","rmse_mean_by_model_fields.png"), p_mean, width = 9, height = 5, dpi = 300)
```

# plot 2

```{r}

fill_min <- floor(min(rmse_summary$mean, na.rm = TRUE))
fill_max <- ceiling(max(rmse_summary$mean, na.rm = TRUE))

p_clear <- ggplot(rmse_summary, aes(x = field_f, y = Model, fill = mean)) +
  geom_tile(color = "white", linewidth = 0.7) +
  geom_text(aes(label = sprintf("%.1f", mean)),
            size = 3.5, fontface = "bold", color = "black") +
  scale_fill_gradient(
    name = "Mean RMSE",
    low = "#f7fbff",  
    high = "#6baed6", 
    limits = c(fill_min, fill_max)
  ) +
  coord_equal() +
  labs(x = "# fields", y = "model") +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    legend.position = "right"
    #plot.margin = margin(10, 10, 10, 10)
  )

p_clear

# ggsave(here::here("results","rmse_mean_graph.png"), p_clear, width = 7.5, height = 4.5, dpi = 300)

```

# table

```{r}

# ------
rmse_summary %>%
  arrange(Model, field_f) %>%
  mutate(`mean (sd)` = sprintf("%.2f (%.2f)", mean, sd)) %>%
  select(Model, field_f, n, `mean (sd)`, median, lower, upper) %>%
  print(n = Inf)

```