### 2025-08-29 codes/R/Add_b1_b2_to_data_2nd_stage.qmd
Add b1 and b2 to the data_2nd_stage

---
title: "Add b1 & b2 to the data_2nd_stage"
author: "Your Name Here"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

# Load data
```{r setup_not_remove, echo=FALSE, include = F}

library(here)

data_2nd_stage <- readRDS(here("data", "data_20230504", "data_2nd_stage.rds"))

```

# Add b1 & b2 into data_2nd_stage
```{r setup_not_remove, echo=FALSE, include = F}
library(data.table)

reg_data  <- raw_sim_data$reg_data[[1]]
# flatten the list-column `data` inside `reg_data`
reg_data_flat <- rbindlist(
  lapply(seq_len(nrow(reg_data)), function(i) {
    df <- reg_data$data[[i]]
    df$sim <- reg_data$sim[i]
    df
  }),
  use.names = TRUE, fill = TRUE
)

# keep only needed columns for the join
reg_data_subset <- reg_data_flat[, .(sim, aunit_id, b1, b2)]

# set as data.table 
setDT(data_2nd_stage)

# merge by sim and aunit_id
data_2nd_stage <- merge(
  data_2nd_stage,
  reg_data_subset,
  by = c("sim", "aunit_id"),
  all.x = TRUE,
  sort = FALSE
)

# save
saveRDS(
  data_2nd_stage,
  here("data", "data_20230504", "data_2nd_stage.rds")
)

```