### 2025-10-08 codes/R/Corner_sol_CF.qmd
CF corner solution

---
title: "CF corner solution"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

## Packages and functions
```{r setup_not_remove, echo=FALSE, include = F}
#--- packages ---#
library(data.table)
library(here)
library(tidyr)
```

## load data and find corner solutions
```{r}

# raw sim data
# raw_sim_data <- readRDS(here("data", "raw_sim_data.rds"))

# List of field settings you ran
n_fields_list <- c(1, 3, 5, 10, 20)

# Function to process one n_fields case
process_field_setting <- function(n_fields) {

  # train/test splits for this field count
  splits_file <- here("data", "train_test_split",
                      sprintf("train_test_splits_%dfields.csv", n_fields))
  splits <- fread(splits_file)

  # estimated EONR results for this field count
  eonr_dir <- here("results", sprintf("CF_outcome_%d_fields", n_fields))
  files <- list.files(eonr_dir, pattern = "_eonr\\.rds$", full.names = TRUE)
  eonr_all <- rbindlist(lapply(files, readRDS))

  # Loop on all test sims
  corner_results <- lapply(1:nrow(splits), function(i) {

    test_id <- splits$test_id[i]
    train_ids <- as.integer(unlist(splits[i, grep("^train_", names(splits)), with = FALSE]))

    # get all train sims used for this test (1, 3, 5, 10, 20)
    train_data <- raw_sim_data$reg_data[[1]] %>%
      .[sim %in% train_ids, ] %>%
      .[, .(sim, data)] %>%
      tidyr::unnest(cols = c(data)) %>%
      as.data.table()

    N_min <- min(train_data$N)
    N_max <- max(train_data$N)

    # get estimated opt_N_hat for this test sim
    eonr_sim <- eonr_all[sim == test_id]

    #  corner solutions
    eonr_sim[, corner := (opt_N_hat == N_min | opt_N_hat == N_max)]
    eonr_sim[, n_fields := n_fields]

    return(eonr_sim)
  })

  return(rbindlist(corner_results))
}

# run for all field counts
corner_all <- rbindlist(lapply(n_fields_list, process_field_setting))

# summarize: % of corner solutions by n_fields
corner_summary <- corner_all[, .(
  n_total = .N,
  n_corners = sum(corner, na.rm = TRUE)
), by = n_fields]

corner_summary

# save 
# saveRDS(corner_summary,
#         here("results", "corner_summary.rds"))

```