### 2025-11-16 Proposal_presentation/Chapters_1&2.qmd
proposal presentation

---
title: "Chapter 1: A new model selection approach based on local economically optimal input rate"
author:
  - name: Taro Mieno$^*$
  - name: Mona Mousavi$^*$
  - name: David Bullock$^\dagger$
institute:
  - $^*$ University of Nebraska Lincoln
  - $^\dagger$ University of Illinois Urbana Champaign 
date: 06/12/2025
format:
  beamer:
    theme: Singapore
    navigation: horizontal
    section-titles: false
    slide-level: 2
    fontsize: "10pt"
header-includes: 
- \logo{\ifnum\thepage>1\includegraphics[width=1cm]{nebraska-n.jpg}\fi}
---

```{r, include = FALSE}
library(terra)
library(sf)
library(tidyterra)
library(ggplot2)
library(data.table)
library(dplyr)
library(flextable)
```

# Background and Research Objectives

## Background

+ Uniform input management strategies can lead to both economic losses and environmental inefficiencies.

+ Variable-rate input management as a solution (accounts for spatial heterogeneity in crop input demand.

+ On-farm precision experiments (OFPE) as a promising way (GPS technologies and variable-rate applicators).
  + Run on-farm agronomic trials and collect data
  + \textcolor{blue}{Estimate how yield responds to nitrogen based on observed characteristics}
  + Identify site-specific optimal nitrogen rates

## Motivation

+ Machine learning approaches being increasingly used for estimating site-specific yield response functions
  + identify the best model among several candidate ML approaches and also tuning hyper parameters via cross-validation as the key part of such process.

+ In all of the previous studies, the best model was selected based on their abilities to predict \textcolor{blue}{yield level} at the observation points (yield-based model selection).

+ However, the ultimate goal is not to predict yield levels, but to estimate yield response (causal marginal impact of nitrogen on crop yield).

## Research Objective

+ Proposes an alternative to conventional yield-based model selection. Specifically, we select models based on their ability to predict local EONR using spatial cross-validation.

+ Evaluate this approach, through Monte Carlo simulations, and compare its performance against the standard yield-based selection method.

# Method

Monte Carlo simulations with 500 iterations.

Steps for a single iteration:

  + Step 1: Generate OFPE experiment data
  + Step 2: Identify the true best (most profitable VRA application rate) model among the five candidate models
  + Step 3: Select the best model among the 5 candidate models based on
    + yield (conventional approach)
    + local-EONR (our approach)
  + Step 4: Evaluate the model selection approaches

The five candidate models:
linear model, linear spatial error (SE) model, random forest (RF), boosted regression forest (BRF), and causal forest (CF).

## Data generation

+ The data generated from a simulated corn production field (about 37.3 hectares) with spatially varying field characteristics.

+ One possible scenario of the spatial variability in the field characteristics is shown.

+ This study assumed the yield data were collected at the subplot level (total 1440 subplots).

## The simulated OFPE field:
```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "95%"
knitr::include_graphics("~/Desktop/g_field_map.png")
```

## True yield responses

The true yield is assumed to follow the quadratic plateau functional form as follows:

$$
y_{j, i}=f_{j, i}(N)= \begin{cases}\alpha_{j, i}+\beta_{j, i} N+\gamma_{j, i} N^2+\varepsilon_{j, i}, & N<\tau_{j, i} \\ \eta_{j,i}+\varepsilon_{j, i}, & N \geq \tau_{j, i}\end{cases}
$$

+ $j \in{1,\ 2,\ 3,...,1440}$: the subplot,
+ $i \in{1,\ ...,6}$: the cell within a subplot,
+ $y_{j,i}$: Corn yield,
+ $N$: Nitrogen rate,
+ $\alpha, \beta, \gamma$: parameters (soil/field characteristics) that vary within a field in a spatially correlated manner and govern how yield responds to $N$
+ $\tau_{j, i}$:  the $N$ rate at which the yield reaches a plateau.
+ $\eta_{j,i}$: the yield at $N = \tau_{j, i}$,
+ $\varepsilon_{j, i}$: spatial autocorrelation.

---

The true economically optimal N rate for each cell is then determined by:

$$
N_{j, i}^* \equiv \operatorname{argmax}_N\left[p \cdot f_{j, i}(N)-w \cdot N\right]
$$

+ $p$: the crop price, 
+ $w$ the nitrogen fertilizer price, 
+ $f_{j, i}(N)$: previously defined true yield response function for the specific cell.

---

## On-farm trial (nitrogen rate)

+ six N trial rates ($\{80, 130, 154, 184, 219, 270\}$ kg/ha.) followed a 6 × 6 Latin square design.

(Note: 8 blocks, each block including 36 plots ($8*36=288$ plots) to establish the Latin square design).

---

```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "100%"
knitr::include_graphics("g_N.png")
```

---

## Yield data

Based on the response parameters $\left(\alpha_{j, i}, \beta_{j, i}, \gamma_{j, i}, \tau_{j, i}, \eta_{j,i}\right)$ and error term $\varepsilon_{j,i}$, the yield $y_{j, i}$ for each cell $(j, i)$ was generated following the true response function.

* Note:
The harvesting equipment was assumed to be able to accurately collect yield data only at the subplot resolution. Therefore, the observed yield data were at the resolution of subplot level.

---

```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "100%"
knitr::include_graphics("/Users/monamousavi/Dropbox/Causal_climate/g_y.png")
```

## independent variables

+ The true response parameters $\left(\alpha_{j, i}, \beta_{j, i}, \gamma_{j, i}, \tau_{j, i}, \eta_{j,i}\right)$ and economically optimal N rates $N_{j, i}^*$ are not directly observable by producers or researchers.

+ This study creates some covariates (correlated with those true parameters) to serve as the observable field characteristics:

$$X_j = \left(\beta^{1,1}_j, \beta^{1,2}_j, \beta^{2,1}_j, \beta^{2,2}_j, \tau^{1,1}_j, \tau^{1,2}_j, \tau^{2,1}_j, \tau^{2,2}_j, \eta^{1,1}_j, \eta^{1,2}_j, \eta^{2,1}_j, \eta^{2,2}_j \right)$$

$$\beta = \beta^1 + \beta^2, \;\;
\beta^1 = \beta^{1,1} + \beta^{1,2}, \;\;\beta^2 = \beta^{2,1} + \beta^{2,2}$$ 
$$\tau = \tau^1 + \tau^2, \;\;
\tau^1 = \tau^{1,1} + \tau^{1,2}, \;\;\tau^2 = \tau^{2,1} + \tau^{2,2}$$ 
$$
\eta = \min(\eta^1, \eta^2), \;\;
\eta^1 = \min(\eta^{1,1}, \eta^{1,2}), \;\;
\eta^2 = \min(\eta^{2,1}, \eta^{2,2})
$$

# Identifying the true best model among the five candidate models

+ Each of the candidate models are trained using the whole experiment data and then variable rate EONR are estimated for the whole field.

+ The estimated variable rate N recommendations are evaluated against the true variable rate N to evaluate the performance of the models (using profit loss relative to the true site-specific EONR as the criteria):

---

The profit loss of model $m$:

$$
\Delta\pi_m = \pi_{true} - \pi_m
$$

Where:

$$
\pi_{true} = \frac{1}{N} \cdot \sum_{i=1}^{N} \left[P_c \cdot f(N^*_i, X_i) - P_N \cdot N^*_i\right]
$$

$$
\pi_{m^*} = \frac{1}{N} \cdot \sum_{i=1}^{N} \left[P_c \cdot f(\widehat{N}^*_{m}(X_i), X_i) - P_N \cdot \widehat{N}^*_{m}(X_i)\right]
$$

## Model Selection

+ For both approaches (yield prediction accuracy, local EONR prediction accuracy) spatial cross-validation is employed.
  + The entire dataset is randomly split into a train and test dataset in a spatially clustered manner, which is repeated to create multiple folds.

## An example of the train-test split for each of the five folds

```{r}
knitr::include_graphics("g_split_example.png")
```

## Model selection based on yield level prediction

For each fold:

1. train the model using the train data of the fold
2. predict yield for the test data based on the trained model ($\hat{y}_{i,f}$)
3. calculate RMSE of yield prediction as 
$$
RMSE\text{-}Yield_{m,f} = \sqrt{\frac{\sum_{i=1}^{N_t}(\hat{y}_{i,f} - y_{i,f})^2}{N_t}}
$$

## Model selection based on local EONR

+ True EONRs are never observable in the real world

+ Using estimated “local uniform EONR” as the proxy for the true local EONR  to overcome this challenge

+ Steps for each fold and candidate models:
  1. train the GAM model using the **test** data 
  2. estimate uniform EONR for the **test** data ($\hat{N}^*_{f,gam}$)
  3. train the candidate model using the **train** data 
  4. estimate site-specific EONR for the **test** data based on the trained candidate model ($\hat{N}^*_{i, f,m}$). 
  5. Calcurate the average of the site-specific EONR ($\hat{N}^*_{f,m}$)
  6. Calcurate the RMSE of the estimated local EONR:

$$
RMSE\text{-}LE_{m} = \sqrt{\frac{\sum_{f=1}^{F} (\hat{N}^*_{f,m} - \hat{N}^*_{f,gam})^2}{F}}
$$

## Evaluation of model selection performance

+ Selected models (based on both approaches) are evaluated based on the following difference:

$$
\pi_{m^*} - \pi_{m^s}
$$

+ $\pi_{m^*}$: the highest profit that could be achieved if (EONR from) the true best model were used. 
+ $\pi_{m^s}$: the profit achieved by the selected model

# Results and Conclusions

## Simulation Results: Performance of the model selection approaches

```{r}
#| fig-align: "center"
#| out-width: "90%"
knitr::include_graphics("g_ranking.png")
```

## Simulation Results: Accuracy of the model selection approaches

```{r}
#| echo: false
#| results: 'asis'
tbl <- readRDS("main_selection_accuracy_table.rds")
invisible(flextable::save_as_image(tbl, path = "table_accuracy.png"))
knitr::include_graphics("table_accuracy.png")

```

## Simulation Results: profit loss associated with the seleted models relative to the actual best models

```{r}
#| fig-align: "center"
#| out-width: "90%"
knitr::include_graphics("g_pi_loss.png")
```

## Simulation Results: Accuracy of Local EONR estimation with GAM

```{r}
#| fig-align: "center"
#| out-width: "90%"
knitr::include_graphics("g_gam_leonr.png")
```

## Conclusion

This study demonstrates the economic advantage of local EONR method in making more accurate and cost-effective input management decisions compared to the yield-based model selection.

## Chapter two:

### Causal methods for Site-specific yield response function estimation

# Background and Research Objectives

## Background

+ Efficient nitrogen management remains one of the most critical challenges in modern agricultural production.

+ Applying right amount of nitrogen at right place (and right time) recognizing the need for nitrogen based on observed characteristics (soil N, soil and field characteristics, weather).

+ An increasing amount of data from on-farm field trials has been accumulated over the years.

+ Estimating yield response functions using data pooled from multiple fields can provide significant agronomic and economic insights.

## Motivation

+ When on-farm trial data is pooled, the treatment variable becomes "continuous" rather than discrete.

+ Existing approaches—such as direct applications of random forests or neural networks—often tend to fail to capture heterogeneous treatment effects of a continuous variable.

+ Most causal machine learning methods (e.g., R-learner) are designed for binary treatments, whereas our objective is to recover the full yield response \textcolor{blue}{curve}.

## Research Objective

+ Find a causal machine learning approach capable of effectively identifying heterogeneous treatment effects tailored for yield response function estimation using data from on-farm trials, where the treatment variable is continuous and randomized.

+ Evaluate the performance of the proposed method through Monte Carlo simulations, comparing it against commonly used existing approaches.

## Method

### Double-ML (binary treatment ($T$) case):

DML works fine in a \textcolor{blue}{binary} treatment ($T$) case (model assumes linearity in $T$):

```{=tex}
\begin{align*}
\mbox{Model 1: } Y = \theta(X)\cdot T + g(X, W) + v
\end{align*}
```

+ $Y$: Dependent variable 
+ $T$: Binary treatment variable (variable of interest) 
+ $X$: Set of variables that affect the impact of $T$ and also $Y$ directly
+ $\theta(X)$: Treatment effect as a function of $X$
+ $W$: Set of variables that affect $Y$ directly
+ $v$: Error term

---

$$
\begin{aligned} & Y_{res}=Y-E[Y \mid X, W]=Y-f(X, W) \\ & T_{res}=T-E[T \mid X, W]=T-h(X, W)\end{aligned}
$$

+ A simple regression of $Y_{res}$ on $T_{res}$ can recover $\theta(X)$:

$$
\begin{aligned}
& Y_{res}= \\
& \begin{aligned}
\theta(X) \cdot T + g(X, W) + v 
- & \underbrace{E\{[\theta(x) \cdot T + g(X, W) + v] \mid X, W\}}= \\
& \textcolor{blue}{=} \textcolor{blue}{\theta(X) \cdot E(T \mid X, W) + g(X, W)} \\
& \textcolor{blue}{=}  \textcolor{blue}{\theta(X) \cdot h(X, W) + g(X, W)}
\end{aligned} \\
& \theta(X) \cdot T + g(X, W) + v 
- \theta(X) \cdot h(X, W) - g(X, W)= \\
& \theta(X) \cdot [\underbrace{T - h(X, W)}_{T_{res}}] + v \\
& \Rightarrow Y_{res} = \theta(X) \cdot T_{res} + v
\end{aligned}
$$

---

### Double-ML (continuous treatment ($T$) case):

DML fails in a \textcolor{blue}{continuous} treatment ($T$) case:

```{=tex}
\begin{align*}
\mbox{Model 2: } Y = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ $T$: <span style = "color: blue;">Continuous </span> treatment variable (e.g., nitrogen rate) 
+ $\theta(X, T)$: Impact of $X$ and $T$ with their full interactions allowed

---

$\begin{aligned}
& Y_{res}= \\
& \theta(X, T)+g(X, W)+v-E\{\theta(X, T)+g(X, W)+v \mid X, W\}= \\
& \theta(X, T)+g(X, W)+v-E(\theta(X, T) \mid X, W)-g(X, W) \\
& \Rightarrow Y_{res}=\theta(X, T)-E(\theta(X, T) \mid X, W)+v
\end{aligned}$

So, orthogonalizing $T$ does not work: $\widehat{\theta}(X, T - g(X, W))$ cannot be used to approximate $E[\theta(X, T)]$.

## Our proposed approaches

+ Single-orthogonalization
+ Extended CF

## Single-orthogonalization

```{=tex}
\begin{align*}
\mbox{Model 2: } Y = \theta(X, T) + g(X, W) + v
\end{align*}
```

1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
2. Regress $Y-f(X,W)$ on $T$

This approach works because The orthogonalization shifts the attention from the variations in $Y$ to the variations in $Y-f(X,W)$ just due to $X$ and $T$.

## Extended CF

+ We propose the use of smoothing splines. In spline regression, a single variable can be decomposed into multiple basis functions in a linear-in-parameter manner.

```{=latex}
\begin{align} \label{eq-spline-cf}
y = \sum_{k=1}^K \theta_k(X)\cdot S_k(T)+ g(X, W) + v
\end{align}
```

+ $S_k(T)$: the value of the basis function $S_k(\cdot)$ evaluated at $T$.

## Models

$$
y = f(X, N) + g(X,W) + v
$$

+ $Y$: Crop yield
+ $N$: Nitrogen rate (lb/acre) 
+ $X$: Collection of variables ($\alpha$, $\beta$, $\gamma$, and $\tau$)
+ $v$: Error term

## Scenarios

+ Estimation approaches:

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("f_est.png")
```

+ For each of 500 simulation rounds, one field is selected as the test field. The remaining fields are combined to construct the training set in five scenarios: 
  + number of fields from which data is combined: 1, 3, 5, 10 and 20

## Site-specific EONR estimation

For a given price of corn and nitrogen EONR is found by solving the following profit maximization problem:

$$
\widehat{\mathrm{EONR}}=\arg \max _N\left[P_{\text {Corn }} \times \hat{f}(X, N)-P_N \times N\right]
$$

+ $\widehat{f}$: yield response curve estimated by the model under evaluation

## Model evaluation

Model evaluation is based on the accuracy of predicting <span style = "color: blue;"> site-specific economically optimal nitrogen rates </span>($\widehat{EONR}$).

$$
RMSE_{EONR}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{EONR}_i-EONR_i\right)^2}
$$

# Results and Conclusions

## Simulation (preliminary) results

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("fig_rmse_hist.png")
```

## Simulation (preliminary) results

```{r}
#| fig-align: "center"
#| out-width: "45%"
knitr::include_graphics("tbl_best_model_by_field_count.png")
```

## Simulation (preliminary) results

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("model_improvement.png")
```

## Conclusion

By combining orthogonalization with flexible final-stage learners, we demonstrate how to preserve the strengths of causal inference while accommodating continuous treatments and agronomic heterogeneity.