### 2025-11-17 Writing/manuscript_CF_continuous.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{anyfontsize}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Estimation of the Effect of a Continuous Treatment Variable},
  pdfauthor={Mona Mousavi},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Estimation of the Effect of a Continuous Treatment Variable}
\author{Mona Mousavi}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Efficient nitrogen (N) management remains one of the most critical
challenges in modern agricultural production. Nitrogen is essential for
achieving high crop yields, yet excessive or poorly timed applications
can lead to economic losses and environmental harm. The principle of
applying the right amount of nitrogen at the right place and time
requires understanding the spatial and temporal variability in crop
nitrogen needs. This variability arises from complex interactions among
soil nitrogen availability, soil and field characteristics, and weather
conditions. Accurately identifying the optimal nitrogen rate for each
site and season is therefore fundamental to improving both profitability
and sustainability in crop production systems.

On-farm field experiments have emerged as a powerful approach to
generate the data necessary for site-specific nitrogen management.
Through on-farm agronomic trials, researchers can directly observe yield
responses to varying nitrogen rates under real-world field conditions.
Analyzing these data enables estimation of yield response functions and
identification of site-specific economically optimal nitrogen rates.
Recent advances in data analytics---particularly the widespread adoption
of machine learning (ML) methods---have expanded the potential of
on-farm experimentation. Two major methodological directions have
emerged: (1) yield prediction--based methods, which recommend
site-specific input rates by predicting yield as a function of nitrogen
and environmental covariates (Krause et al. (2020); Barbosa et al.
(2020); Barbosa, Hovakimyan, and Martin (2020); Gardner, Mieno, and
Bullock (2021)), and (2) causal machine learning (causal ML) methods
(Kakimoto et al. (2022)), which estimate the underlying causal
relationship between nitrogen application and yield response.

In the OFPE setting, causal machine learning methods have shown to be
promising to recover the full yield response curve. Kakimoto et al.
(2022) showed that causal forest can estimate heterogeneous treatment
effects of an input on yield more accurately than prediction-oriented
machined approaches like random forest, boosted forest, and
convolutional neural network. However, their application of CF is only
limited to a single-year OFPE experiment from a single field. Causal
forest was developped primarily to estimate the heterogeneous treatment
effects of a dicrete treatment. In OFPE, there are typically five or six
distinctive target experimental rates (see the first panel of
Figure~\ref{fig-distribution-N}). In Kakimoto et al. (2022), CF
estimates the effect of multiple treatments defined as the increase in
nitrogen rate from the lowest rate to the other rates (so n − 1
tretments in n-rate trials). However, once data from multiple fields
with different target rates are combined, applied nitrogen rates become
rather continuous as seen in the second panel of
Figure~\ref{fig-distribution-N}. Existing approaches---such as direct
applications of random forests or neural networks and CF---often tend to
fail to capture heterogeneous treatment effects of a continuous variable
(Kakimoto et al. (2022)).

\begin{figure}

{\centering \includegraphics{Figures/g_N_hist_single_multiple.pdf}

}

\caption{\label{fig-distribution-N}Distribution of nitrogen rates from a
single OFPE and multiple OFPEs}

\end{figure}

As multi-year, multi-field OFPE datasets accumulate, pooling data across
fields has become both feasible and desirable---it can improve
estimation of yield response functions and deliver stronger agronomic
and economic insights. It creates a need for methods that handle
continuous treatments and it becomes especially important to adapt
causal methods to continuous treatments without losing their statistical
strengths over other machine-learning approaches.

The key objective of this article is to identify a causal machine
learning approach capable of estimating heterogeneous treatment effects
for yield response function estimation using data from on-farm trials,
where the treatment variable---nitrogen rate---is continuous and
randomized. We investigate two complementary approaches: a
Single-Orthogonalized Neural Network (SO-ANN), designed to flexibly
capture nonlinear treatment--response relationships, and an extended
Causal Forest (CF) that incorporates smoothing splines to model
continuous treatments. Both methods aim to preserve the strengths of
causal ML---nonparametric modeling of complex interactions without
restrictive assumptions. Their performance is evaluated through Monte
Carlo simulations that mimic typical OFPE conditions and crop yield
responses to nitrogen.

We find that the SO-ANN approach outperforms the non-orthogonalized
model, which represents the common practice in yield response function
estimation. In contrast, the Double-Orthogonalized approach performs
poorly, and its accuracy does not improve even with larger sample sizes,
consistent with theoretical expectations. Our extended CF method
successfully captures heterogeneous and nonlinear effects of a
continuous treatment and accurately estimates site-specific
profit-maximizing nitrogen rates with minimal bias underscoring the
promise of this approach. Although the methodological extension proposed
here is primarily motivated by the needs of OFPE, it holds potential for
a wide range of applications in applied economics where continuous
treatments and heterogeneous effects are of interest.

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{models-and-estimation-procedures}{%
\subsection{Models and estimation
procedures}\label{models-and-estimation-procedures}}

The standard approach in the precision-agriculture literature for
estimating site-specific input response is to model the outcome level
directly as a function of the treatment/input and covariates---e.g.,
estimate \(Y=f(I,C)\) where \(I\) is the input (like nitrogen rate) and
\(C\) are field characteristics---and then derive optimal input
decisions from that estimated function. This is what is typically done
in the recent on-farm precision experimentation literature (e.g.,
Barbosa, Hovakimyan, and Martin (2020); Barbosa et al. (2020); Krause et
al. (2020); Gardner, Mieno, and Bullock (2021)). The problem is that
these methods optimize prediction of yield levels---not the causal
effect of changing the input. Good level prediction at the observed
input does not guarantee accurate recovery of the response function
needed for optimal input choice, because the fitted \(f(I,C)\) can
conflate input variation with confounding and misattribute the drivers
of yield variation.

To identify causal effects rather than just predict levels in the single
binary treatment case, researchers use double/debiased machine learning
(DML), which introduces orthogonalization so the treatment-effect
estimator is insensitive to small errors in the auxiliary fits
(Chernozhukov et al. (2018); Mackey, Syrgkanis, and Zadik (2018)).
Consider the structural model (Model 1):

\[
Y=\theta(X) T+g(X, W)+v
\]

where \(Y\) is the outcome, \(T \in\{0,1\}\) is the treatment, \(X\)
covariates that also modify the treatment effect \(\theta(X)\), \(W\)
additional predictors of \(Y\), and \(v\) is noise. Orthogonalization
estimates \(f(X, W)=\mathbb{E}[Y \mid X, W]\) and
\(h(X, W)=\mathbb{E}[T \mid X, W]\), forms residuals
\(Y_{\mathrm{res}}=Y-f(X,W)\) and \(T_{\mathrm{res}}=T-h(X, W)\), and
then relates \(Y_{\mathrm{res}}\) to \(T_{\mathrm{res}}\). Because this
score is Neyman-orthogonal and is typically implemented with
cross-fitting, small errors in \(f\) and \(h\) do not induce first-order
bias in \(\theta(X)\).

The identifying algebra in Model 1 follows from linearity in \(T\):

\[
\begin{align*}
\mathbb{E}[Y \mid X, W]=\theta(X) \mathbb{E}[T \mid X, W]+g(X, W)=\theta(X) h(X, W)+g(X, W)
\end{align*}
\]

so

\[
\begin{align*}
Y_{\mathrm{res}}=Y-\mathbb{E}[Y \mid X, W]=\theta(X)(T-h(X, W))+v=\theta(X) T_{\mathrm{res}}+v
\end{align*}
\]

Thus regressing \(Y_{\mathrm{res}}\) on \(T_{\mathrm{res}}\) recovers
\(\theta(X)\) up to noise.

Kakimoto et al. (2022) applied this framework to estimate heterogeneous
treatment effects of nitrogen on yield and treated heterogeneity in a
discrete way. However, when data are pooled across fields and applied
nitrogen rates become continuous, the standard residualization of the
treatment in DML does not extend cleanly to that setting.

Consider the continuous-treatment setting, where the structural model is

\[
\begin{align*}
Y=\theta(X, T)+g(X, W)+v
\end{align*}
\]

with \(T\) (e.g., nitrogen rate) continuous, and \(X\)
heterogeneity-inducing covariates that interact with \(T\) through the
flexible response function \(\theta(X, T)\). In this case
double-orthogonalization fails because \(\theta\) is a general
(nonlinear and interaction-rich) function of \(T\). Residualizing the
treatment by replacing \(T\) with \(T-\mathbb{E}[T \mid X, W]\) does not
produce a valid adjustment for the heterogeneous effect. Formally,

\[
\begin{align*}
\theta(X, T)-\mathbb{E}[\theta(X, T) \mid X, W] \neq \theta(X, T-\mathbb{E}[T \mid X, W])
\end{align*}
\]

and the clean decomposition that underlies double-orthogonalization in
the linear-in-\(T\) case breaks down. Instead, we apply
single-orthogonalization (SO). We first estimate
\(f(X, W)=\mathbb{E}[Y \mid X, W]\) and form the residual
\(Y_{\mathrm{res}}=Y-f(X, W)\), and then recover \(\theta(X, T)\) by
flexibly modeling \(Y_{\mathrm{res}}\) as a function of \((X,T)\). This
approach removes baseline variation due to \(g(X,W)\) while preserving
the full interaction structure between \(X\) and the continuous
treatment.

To study and compare these ideas systematically, we adopt a single
underlying structural model and then implement alternative
orthogonalization schemes.

The model of interest throughout is (Model 3)

\[
\begin{align*}
y = f(X, N) + g(W) + v
\end{align*}
\]

where \(y\) is crop yield, \(N\) is the continuous nitrogen rate, \(X\)
is a set of covariates that induce heterogeneity in the impact of \(N\)
on \(y\), and \(W\) is a set of covariates that affect \(y\) but do not
interact with \(N\). We implement three estimation
strategies---double-orthogonalized, single-orthogonalized, and
non-orthogonalized---that differ only in how (and whether) the outcome
and treatment are residualized before the final fit. Orthogonalization
means fitting a variable on its conditioning covariates and using the
residual (observed minus predicted) so that the subsequent learner
focuses on variation not explained by those covariates; all
residualizations are performed with random forests. In the
double-orthogonalized approach, both \(y\) and \(N\) are residualized
with respect to \((X,W)\) and the final learner is trained on those
residuals. In the single-orthogonalized approach, only \(y\) is
residualized and the raw \(N\) is retained. The non-orthogonalized
specification skips residualization entirely and fits the learner
directly to \((y,N,X,W)\). For each strategy we consider three
final-stage learners---neural networks (NN), random forests (RF), and
causal forests (CF)---to evaluate how the orthogonalization scheme and
learning algorithm interact in recovering heterogeneous nitrogen
response.

The table below shows all the cases we consider.

\includegraphics[width=4in,height=\textheight]{f_est.png}

\hypertarget{single-orthogonalized-nn-so-nn-and-so-rf}{%
\subsubsection{Single-orthogonalized NN (SO-NN) and
(SO-RF)}\label{single-orthogonalized-nn-so-nn-and-so-rf}}

In the single-orthogonalized framework only the baseline component
\(g(W)\) is partialed out before the final-stage estimation; the
continuous treatment \(N\) and heterogeneity-inducing covariates \(X\)
remain in their original form. Formally, starting from Model 3, we first
estimate \(g(W)=\mathbb{E}[y \mid W]\) using a cross-fitted random
forest and form the residualized outcome

\[
\begin{align*}
\tilde{y}=y-\widehat{g}(W)
\end{align*}
\]

The second stage fits a flexible model of \(\tilde{y}\) on \((X,N)\) to
recover the heterogeneous response function \(f(X,N)\). The distinction
between SO-NN and SO-RF is only in the choice of the final learner:
SO-NN uses a neural network, while SO-RF uses a random forest. In both
cases the learner estimates \(\hat{f}(X, N)\), which is then used to
predict yield under candidate nitrogen rates. Marginal profit is
computed over a grid of \(N\) values and the economically optimal
nitrogen rate is selected as the one that maximizes

\[
\begin{align*}
\widehat{\mathrm{EONR}}=\arg \max _N\left[P_{\text {Corn }} \times \hat{f}(X, N)-P_N \times N\right]
\end{align*}
\] where \(P_{\text {Corn }}\) is the corn price, \(P_{\text {N }}\) is
the nitrogen price, and \(\hat{f}\) is the estimated yield response.

\hypertarget{double-orthogonalized-cf-do-cf-and-ann-do-ann}{%
\subsubsection{Double-orthogonalized CF (DO-CF) and ANN
(DO-ANN)}\label{double-orthogonalized-cf-do-cf-and-ann-do-ann}}

\hypertarget{docf-continuous}{%
\paragraph{DO‑CF (continuous)}\label{docf-continuous}}

To adapt double-orthogonalization to the model of interest (Model 3), we
estimate the heterogeneous response \(f(X,N)\) by expanding the
continuous treatment \(N\) in a set of spline basis functions and
allowing their coefficients to vary with the heterogeneity-inducing
covariates \(X\):

\[
\begin{align*}
y=\sum_{k=1}^K \theta_k(X) S_k(N)+g(X, W)+v
\end{align*}
\]

Although double-orthogonalization generally fails when \(f(X,N)\) is an
arbitrary nonlinear function of \(N\), this spline-based formulation
restores the key linear-in-parameter structure that makes DO valid,
because the model is linear in the transformed treatment basis
\(\left\{S_k(N)\right\}\). Residualizing each basis component and the
outcome with respect to \((X,W)\) then yields a clean decomposition
analogous to the linear-in-\(T\) case. Let
\(\widetilde{S}_k(N)=S_k(N)-\mathbb{E}\left[S_k(N) \mid X, W\right]\).
Then

\[
\begin{align*}
\tilde{y}=y-\mathbb{E}[y \mid X, W]=\sum_{k=1}^K \theta_k(X) \widetilde{S}_k(N)+v,
\end{align*}
\]

so causal forest can nonparametrically estimate each heterogeneous
weight \(\theta_k(X)\) by relating the residualized outcome to the
residualized spline bases. This Neyman-orthogonal (R-learner) style
decomposition preserves the ability to capture complex, heterogeneous,
and nonlinear treatment effects while inheriting the robustness of
double-orthogonalization: small errors in the nuisance fits for
\(g(X,W)\) or the conditional means of the bases do not induce
first-order bias in \(\theta_k(X)\). The estimated \(\theta_k(X)\) then
combine to recover the full heterogeneous response function and, by
differencing across candidate \(N\) values, site-specific treatment
effects and economically optimal nitrogen rates.

\hypertarget{doann}{%
\paragraph{DO‑ANN}\label{doann}}

In the DO-ANN approach, we apply double-orthogonalization to Model 3 by
residualizing both the outcome \(y\) and the continuous treatment \(N\)
with respect to \((X, W)\) using cross-fitted random forests:

\[
\begin{align*}
\tilde{y}=y-\widehat{\mathbb{E}}[y \mid X, W], \quad \tilde{N}=N-\widehat{\mathbb{E}}[N \mid X, W]
\end{align*}
\]

The final-stage learner is a neural network that models \(\tilde{y}\) as
a linear function of \(\tilde{N}\), with heterogeneity in the slope
captured by covariates \(X\):

\hat{\tilde{y}}=\hat{\theta}(X) \cdot \tilde{N}
\end{align*}

This structure preserves the linear-in-treatment-residual form required
for valid double-orthogonalization while allowing \(\hat{\theta}(X)\) to
vary flexibly across sites. The neural network receives \(X\) as input
and outputs the site-specific treatment effect.

After training, we reconstruct \(\hat{f}(X, N)\) over a grid of
candidate nitrogen rates by evaluating the model across values of \(N\),
and select the economically optimal nitrogen rate.

\hypertarget{non-orthogonalized-ann-no-nn-and-rf-no-rf}{%
\subsubsection{Non-orthogonalized ANN (NO-NN) and RF
(NO-RF)}\label{non-orthogonalized-ann-no-nn-and-rf-no-rf}}

In the non‑orthogonalized setting, neither the outcome variable nor the
treatment is residualized prior to estimation. Instead, the model
directly fits the observed yield on both the nitrogen rate and
covariates. We estimate the response function \(\hat{Y}=\hat{f}(X, N)\)
using the original values of \(Y\), \(N\) and \(X\), without any
orthogonalization. As in the single‑orthogonalized approaches, the
function \(\hat{f}(X, N)\) is estimated non‑parametrically. The
difference between NO‑NN and NO‑RF lies in the choice of final-stage
learner: NO‑NN uses a neural network, while NO‑RF relies on a random
forest. Once the yield response function is estimated, the EONR is
determined using the same profit-maximization procedure described in
earlier models.

\hypertarget{performance-measurement}{%
\subsection{Performance Measurement}\label{performance-measurement}}

For each simulation round, we evaluate model performance by calculating
the root mean squared error (RMSE) of EONR predictions and the
associated profit loss for each method. The RMSE of EONR predictions in
a given simulation round is computed as follows:

\begin{align*}
RMSE_{EONR}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{EONR}_i-EONR_i\right)^2}
\end{align*}

To assess the predictive accuracy and economic relevance of these
modeling approaches, we design a set of simulation experiments that
reflect realistic agronomic conditions. These simulations serve as a
controlled environment for evaluating model performance in estimating
yield response functions and recommending optimal nitrogen rates.

\hypertarget{simulation-scenarios}{%
\subsection{Simulation Scenarios}\label{simulation-scenarios}}

To evaluate the performance of the proposed models we design Monte Carlo
simulations that mimic a typical on-farm precision experimentation
(OFPE) setting. These simulations examine the ability of each method to
accurately estimate the site-specific yield response to nitrogen
application and the EONR under varying data availability conditions.

Our simulation consists of 500 independently generated fields, each
representing a distinct agronomic environment with field-specific
parameters that shape the crop's response to nitrogen application. These
parameters vary systematically across fields to introduce heterogeneity
in yield responses. The true yield response in each field follows a
quadratic-plateau functional form, a widely used structure for modeling
nitrogen response curves in agronomic studies.

For each simulation round, one field is selected as the test field. The
remaining fields are used to construct the training set in three
scenarios: (1) one-field training, (2) three-field training, (3)
five-field training, (4) ten-field training and (5) twenty-field
training. These scenarios allow us to assess how model performance
changes with the size and diversity of training data. Within each round,
the model is trained on the designated number of training fields and
then used to estimate the yield response function and the site-specific
EONR for the test field.

The experimental design within each field consists of randomized
nitrogen treatment rates applied across plots. The nitrogen rates vary
between fields and are drawn to reflect realistic trial setups that span
a range of agronomic optima. For evaluation purposes, we use a grid of
candidate nitrogen rates to compute predicted yield and associated
marginal profit for each observation. The predicted EONR is the nitrogen
rate that maximizes predicted profit.

\hypertarget{data-generation}{%
\subsection{Data generation}\label{data-generation}}

First, a 92.17-acre rectangular field is created, which consists of
\(384\) of 60 feet \(\times\) 240 feet plots. each of which will be
assigned an N fertilizer application rate. Each plot is made up of four
(4-rows \(\times\) 1-column) 60 feet \(\times\) 60 feet ``subplots,''
which are the unit of analysis used in subsequent statistical analysis.
Each subplot consists of a 6-rows \(\times\) 6-columns grid of
thirty-six 10 feet \(\times\) 10 feet ``cells.'' Visual representation
of the field layout is presented in Figure~\ref{fig-field-layout}. The
data are generated at the cell-level first and then aggregated up to
analysis-unit level.

\begin{figure}

{\centering \includegraphics{Figures/g_layout.pdf}

}

\caption{\label{fig-field-layout}Visualization of the field layout and
components}

\end{figure}

Cell-level yields are generated following the quadratic-plateau
functional form as follows:

\begin{align} 
y_{i,j} = 
  \begin{cases}
  \alpha_{i,j} + \beta_{i,j} N + \gamma_{i,j} N^2 + \varepsilon_{i,j}, & N < \tau_{i,j} \\
  \eta_{i,j} + \varepsilon_{i,j}, & N \geq \tau_{i,j}
  \end{cases}
\label{eq:yield-gen-sim}
\end{align}

In this formulation, yield increases as \(N_{i,j}\) increase until
\(N_{i,j}\) reaches \(\tau_{i,j}\), at which yield hits the plateau,
\(\eta_{i,j}\). The yield level at \(N_{i,j}= 0\) is \(\alpha_{i,j}\).
The rate at which yield increases with respect to \(N_{i,j}\) is
governed by \(\beta_{i,j}\) and \(\gamma_{i,j}\). For a given cell, the
quadratic-plateu function has five parameters, but the degree of freedom
is only three. That is, once any three of the five parameters are
determined, then the rest of the parameters are automatically
determined. Specifically, we first generated \(\alpha_{i,j}\),
\(\eta_{i,j}\), and \(\tau_{i,j}\) in a spatially correlated manner
following the Gaussian bluh bluh process (cite), and then
\(\beta_{i,j}\) and \(\gamma_{i,j}\) were derived. The erorr term
(\(\varepsilon_{i,t}\)) is generated using the same spatial process.
This is repeated independenty 500 times to create 500 fields of the same
size but different spatial distribution and values of the three
parameters.

Trial designs vary field by field based on the value of \(\tau_{i,j}\).
Specifically, the lowest nitrogen rate is set at the minimum of
\(\tau_{i,j}\) of the field minus 50, and the highest rate is set at the
maximum of \(\tau_{i,j}\) of the field plus 30. We then create a
sequence of equi-distance 5-level nitrogen rates. Once the unique
nitrogen levels are determined, a Latin square design is used as a trial
design (see Figure~\ref{fig-trial-design}) for an example trial design).
This ensures that nitrogen rate is orthogonal to any variable including
the error term, avoiding endogeneity problem as done in OFPE in real
world. Now, yield values at the cell level can be calculated by plugging
in the value of \(N_{i,j}\), \(\alpha_{i,j}\), \(\eta_{i,j}\),
\(\tau_{i,j}\), and \(\varepsilon_{i,j}\) into
Equation\textasciitilde{}\ref{eq:yield-gen-sim}.

\begin{figure}

{\centering \includegraphics{manuscript_CF_continuous_files/figure-pdf/fig-trial-design-1.pdf}

}

\caption{\label{fig-trial-design}Example trial design}

\end{figure}

\hypertarget{simulation-results}{%
\section{Simulation Results}\label{simulation-results}}

We compare six estimators---CF, SO\_ANN, DO\_ANN, NO\_ANN, SO\_RF, and
NO\_RF---over 500 Monte-Carlo replications and five training sizes (1,
3, 5, 10 and 20 fields). Each replication contains 1,440 evaluation
units. Accuracy is assessed by the RMSE of the estimated economically
optimal nitrogen rate (EONR) with respect to the true (simulated) EONR.

Taken together, Figure~\ref{fig-rmse-hist} and
Figure~\ref{fig-avg-performance} illustrate clear performance patterns
across methods and training sizes. CF delivers low and tightly
concentrated RMSE distributions even with limited data---consistently
maintaining strong accuracy and stability for 1--5 fields (mean RMSE ≈
14.7, 13.1, and 12.1, respectively), with only modest additional gains
at 10 and 20 fields (11.9--11.8). SO\_ANN initially lags behind CF (18.4
at one field), but its mean RMSE decreases steadily as more fields are
added, surpassing CF performance at 10 and 20 fields (11.2 and 8.5,
respectively). This pattern underscores the neural network's superior
ability to exploit richer spatial and agronomic heterogeneity when
sufficient training data are available.

By contrast, NO\_ANN exhibits substantial instability in the
single-field case, with a heavy right tail and much higher errors, but
rapidly converges toward the leading methods by 5--10 fields (mean RMSE
improving to around 14--13), highlighting its strong dependence on
sample size. SO\_RF improves modestly with more fields (mid-20s RMSE for
10--20 fields) but remains consistently less accurate than SO\_ANN or
CF. NO\_RF maintains broad, right-shifted distributions and higher mean
errors throughout (≈ 36--40), confirming its role as a weak baseline.
Finally, DO\_ANN performs poorly throughout, with both histograms and
means showing persistently high errors (≈ 90+) and negligible benefit
from additional data, reflecting over-correction in the
orthogonalization process.

\begin{figure}

{\centering \includegraphics{manuscript_CF_continuous_files/figure-pdf/fig-rmse-hist-1.pdf}

}

\caption{\label{fig-rmse-hist}RMSE histograms by model and number of
fields.}

\end{figure}

\begin{figure}

{\centering \includegraphics{manuscript_CF_continuous_files/figure-pdf/fig-avg-performance-1.pdf}

}

\caption{\label{fig-avg-performance}Average performance by model and
number of fields.}

\end{figure}

Figure~\ref{fig-improvement} reports each method's percentage
improvement in mean RMSE relative to the NO\_RF baseline. CF
consistently provides the strongest gains among the robust approaches
when data are scarce---reducing error by ≈63\% with a single field, and
climbing to ≈67--68\% once 5--20 fields are available. SO\_ANN exhibits
a similar improvement trend but grows even more dominant as more diverse
data accumulate, achieving ≈53\% improvement at 1 field, ≈63\% by 5
fields, and reaching ≈68\% at 10 fields and ≈77\% at 20 fields, the
highest improvement across all models. NO\_ANN transitions from a large
negative result when trained on just one field (substantially worse than
baseline) to strong relative improvements exceeding 65\% by 10--20
fields, reinforcing its reliance on adequate sample support. SO\_RF
offers moderate improvement (≈38--43\%) that changes little with added
fields. DO\_ANN remains entirely uncompetitive across all sample
sizes---showing worsening performance relative to NO\_RF by more than
100--150\%, indicating severe model misspecification.

Together, these results emphasize that while CF is the most reliable
option under limited data availability, SO\_ANN ultimately yields the
largest relative error reduction when multi-field datasets are
accessible, highlighting its superior scalability with richer training
information.

\begin{figure}

{\centering \includegraphics{manuscript_CF_continuous_files/figure-pdf/fig-improvement-1.pdf}

}

\caption{\label{fig-improvement}Improvement relative to NO\_RF.}

\end{figure}

\textbf{?@tbl-best\_model\_by\_field\_count} highlights the
top-performing model for each training size. CF is the dominant method
when only a few fields are available, delivering the lowest mean RMSE
for 1--5 fields. As field diversity increases, SO\_ANN begins to
outperform CF, becoming the preferred approach at 10 fields and showing
a substantial further accuracy gain at 20 fields.

\begin{table}
\caption*{
{\large Best models by field count} \\ 
{\small Ordered by lowest mean RMSE}
} 
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lcr}
\toprule
Model & \# fields & mean RMSE \\ 
\midrule\addlinespace[2.5pt]
SO\_ANN & 20 fields & 8.545 \\ 
SO\_ANN & 10 fields & 11.233 \\ 
CF & 5 fields & 12.115 \\ 
CF & 3 fields & 13.092 \\ 
CF & 1 field & 14.732 \\ 
\bottomrule
\end{tabular*}
\end{table}

Figure~\ref{fig-yield-res-combined} displays predicted and true yield
response functions for the best-performing models at each training size.
For the CF, shown for 1, 3, and 5 fields, the estimated curves closely
track the overall shape of the true response but exhibit a slight upward
bias in predicted yield levels. The general curvature, including
diminishing returns at higher nitrogen rates, is captured consistently
even under data-constrained conditions. As more fields are incorporated,
the CF predictions become smoother and more aligned with the true
function, reflecting improved generalization.

For the SO\_ANN, shown for 10 and 20 fields, the estimated response
matches the true curve more closely in both slope and curvature,
particularly at higher nitrogen levels where CF tended to flatten. At 20
fields, SO\_ANN produces the most accurate approximation overall, with
prediction and truth nearly coincident across the nitrogen range. These
differences visually reinforce the earlier quantitative results: CF
offers strong robustness with sparse multi-field data, while SO\_ANN
increasingly excels as richer spatial heterogeneity becomes available,
enabling superior recovery of the underlying agronomic yield response
function.

\begin{figure}

{\centering \includegraphics{manuscript_CF_continuous_files/figure-pdf/fig-yield-res-combined-1.pdf}

}

\caption{\label{fig-yield-res-combined}Yield response functions for CF
and SO\_ANN models across different field counts.}

\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

This study evaluates alternative causal-ML strategies for estimating
heterogeneous yield response to a continuous nitrogen treatment in an
OFPE setting. CF consistently attains low and tightly concentrated EONR
errors with only 1--5 training fields, indicating strong small-sample
stability (see Figure~\ref{fig-rmse-hist} and
Figure~\ref{fig-avg-performance}). SO\_ANN improves monotonically with
additional fields and ultimately dominates once ≥10 fields are
available, achieving the lowest mean RMSE overall (e.g., 11.2 at 10
fields and 8.5 at 20 fields). These complementary strengths that CF is
preferred under data-constrained conditions, whereas SO\_ANN becomes the
method of choice as multi-field datasets accumulate (see
\textbf{?@tbl-best\_model\_by\_field\_count}).

Methodologically, the contrast between single-orthogonalization and
double-orthogonalization is central. For continuous treatments, the
double-orthogonal residualization of the treatment does not generally
preserve the relevant heterogeneous response structure. Our simulations
reflect this: DO\_ANN performs poorly and does not benefit from larger
samples. In contrast, single-orthogonalization---residualizing only the
outcome to remove baseline variation from covariates \(W\)---retains the
full interaction between \(X\) and \(N\), allowing the final learner to
recover the nonlinearity and heterogeneity of the response function. The
extended CF strikes a different balance: by expressing \(N\) through a
spline basis and orthogonalizing the basis components, it restores a
linear-in-parameters score that CF can learn reliably.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

By combining orthogonalization with flexible final-stage learners, we
demonstrate how to preserve the strengths of causal inference while
accommodating continuous treatments and agronomic heterogeneity.

Across 500 simulated test fields, two approaches consistently dominate:
the extended causal forest and the single-orthogonalized neural network.
CF is highly robust in small-sample environments, achieving strong
accuracy even with training data from just one to five fields. As the
number of training fields increases, SO\_ANN increasingly improves as a
result of spatial and agronomic diversity, ultimately delivering the
best predictive accuracy of all methods once ten or more fields are
available. In contrast, non-orthogonalized baselines require larger
samples to become competitive and the doubly-orthogonalized ANN performs
poorly in all conditions due to the incompatibility between
double-orthogonalization and continuous treatments. Methodologically,
our results highlight the importance of aligning the orthogonalization
strategy with the structure of the treatment variable in causal ML
applications.

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-barbosa2020risk}{}}%
Barbosa, Alexandre, Naira Hovakimyan, and Nicolas F Martin. 2020.
{``Risk-Averse Optimization of Crop Inputs Using a Deep Ensemble of
Convolutional Neural Networks.''} \emph{Computers and Electronics in
Agriculture} 178: 105785.

\leavevmode\vadjust pre{\hypertarget{ref-barbosa2020modeling}{}}%
Barbosa, Alexandre, Rodrigo Trevisan, Naira Hovakimyan, and Nicolas F
Martin. 2020. {``Modeling Yield Response to Crop Management Using
Convolutional Neural Networks.''} \emph{Computers and Electronics in
Agriculture} 170: 105197.

\leavevmode\vadjust pre{\hypertarget{ref-chernozhukov2018double}{}}%
Chernozhukov, Victor, Denis Chetverikov, Mert Demirer, Esther Duflo,
Christian Hansen, Whitney Newey, and James Robins. 2018.
{``Double/Debiased Machine Learning for Treatment and Structural
Parameters.''} Oxford University Press Oxford, UK.

\leavevmode\vadjust pre{\hypertarget{ref-gardner2021economic}{}}%
Gardner, Grant, Taro Mieno, and David S Bullock. 2021. {``An Economic
Evaluation of Site-Specific Input Application r x Maps: Evaluation
Framework and Case Study.''} \emph{Precision Agriculture} 22: 1304--16.

\leavevmode\vadjust pre{\hypertarget{ref-kakimoto2022causal}{}}%
Kakimoto, Shunkei, Taro Mieno, Takashi ST Tanaka, and David S Bullock.
2022. {``Causal Forest Approach for Site-Specific Input Management via
on-Farm Precision Experimentation.''} \emph{Computers and Electronics in
Agriculture} 199: 107164.

\leavevmode\vadjust pre{\hypertarget{ref-krause2020random}{}}%
Krause, M, Savanna Crossman, Todd DuMond, Rodman Lott, Jason Swede,
Scott Arliss, Ron Robbins, Daniel Ochs, and Michael A Gore. 2020.
{``Random Forest Regression for Optimizing Variable Planting Rates for
Corn and Soybean Using Topographical and Soil Data.''}

\leavevmode\vadjust pre{\hypertarget{ref-mackey2018orthogonal}{}}%
Mackey, Lester, Vasilis Syrgkanis, and Ilias Zadik. 2018. {``Orthogonal
Machine Learning: Power and Limitations.''} In \emph{International
Conference on Machine Learning}, 3375--83. PMLR.

\end{CSLReferences}

\end{document}