### 2025-11-17 codes/R/CF_YRF.qmd

---
title: "CF Loop"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

## Packages and functions
```{r setup_not_remove, echo=FALSE, include = F}
#--- packages ---#
library(tidyverse)
library(data.table)
library(mgcv)
library(spatialsample)
library(parallel)
library(sf)
library(grf)
library(ranger)
library(xgboost)
library(mgcv)
library(here)
library(reticulate)
library(readr)
library(dplyr)
library(tidyr)
library(pasim)

Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  VECLIB_MAXIMUM_THREADS = "1"  # macOS Accelerate safeguard
)

#--- python functions ---#
# use_condaenv("cf_conda311", required = TRUE)
# Apple M2
use_condaenv("cf_conda311",
             conda = "/Users/monamousavi/miniforge3/bin/conda",
             required = TRUE)
py_config()
py_run_string("from econml.dml import CausalForestDML; print('CausalForestDML imported OK')")

#--- R functions ---#
source(here("codes", "R", "Utility.R"))

#--- python functions ---#
source_python(here::here("codes/Python/run_CF_c.py"))
```

## load data

```{r}
#raw_sim_data <- readRDS(here::here("data/raw_sim_data.rds"))

# how many fields to train on?
n_fields <- 20

splits_file <- here(
  "data", "train_test_split",
  sprintf("train_test_splits_%dfields.csv", n_fields)
)
output_dir <- here(
  "results",
  sprintf("CF_outcome_%d_fields", n_fields)
)
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

progress_csv <- file.path(output_dir, "progress.csv")

# ─── LOAD SPLITS ────────────────────────────────────────────────────────────────

train_test_splits <- read_csv(splits_file)

```

# CF YRF

```{r}
# ── EONR subfolder ────────────────────────────────────────────────
eonr_dir <- file.path(output_dir, "eonr")
if (!dir.exists(eonr_dir)) dir.create(eonr_dir, recursive = TRUE)

## ── Loop over test sims ───────────────────────────────────────────────────────
#for (i in seq_len(nrow(train_test_splits)))
for (i in seq_len(1)){
#for (i in seq_len(nrow(train_test_splits))) {
  test_sim_id <- train_test_splits$test_id[i]
  out_fn <- file.path(output_dir, sprintf("sim_%03d.rds", as.integer(test_sim_id)))

  # skip if already saved
  if (file.exists(out_fn)) {
    message("Skipping sim ", test_sim_id, " (already saved)")
    next
  }

  tryCatch(
    {
      ## 1) train/test IDs
      split_row     <- train_test_splits[i, ]
      train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:n_fields)]))
      sim_id_ls     <- c(train_sim_ids, test_sim_id)

      ## 2) train & test data
      data <- raw_sim_data$reg_data[[1]] %>%
        .[sim %in% sim_id_ls, ] %>%
        .[, .(sim, data)] %>%
        tidyr::unnest(cols = c(data)) %>%
        as.data.table()

      train_data <- data[sim != test_sim_id, ]
      test_data  <- raw_sim_data$reg_data[[1]] %>%
        .[sim == test_sim_id, ] %>%
        .[, .(sim, data)] %>%
        tidyr::unnest(cols = c(data)) %>%
        as.data.table()

      ## 3) CF 
      x_vars <- c("Nk", "plateau", "b0")
      T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
      Y      <- train_data[, yield]
      X      <- as.matrix(train_data[, ..x_vars])
      W      <- X
      X_test <- as.matrix(test_data[, ..x_vars])

      ## 4) run cf in python
      Sys.setenv(
        OMP_NUM_THREADS      = "1",
        MKL_NUM_THREADS      = "1",
        KMP_AFFINITY         = "disabled",
        KMP_DUPLICATE_LIB_OK = "TRUE"
      )

      te_hat_cf <- run_CF_c_py(
        Y,
        T_info$T_sp,
        X,
        W,
        n_estimators = 2000
      )
      te_info <- get_te(te_hat_cf, test_data, x_vars, "aunit_id")

      ## 5) Response curves
      T_seq <- train_data[, seq(min(N), max(N), length = 100)]
      response_data <- find_response_semi(T_seq, T_info, te_info)

out_dir <- here::here(
  "results",
  "yield_response_function_for_one_iteration"
)
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

saveRDS(
  list(
    te_hat_cf     = te_hat_cf,
    te_info       = te_info,
    T_seq         = T_seq,
    response_data = response_data,
    test_data     = test_data,
    train_data    = train_data
  ),
  file = file.path(out_dir, sprintf("YRF_CF_%dfields.rds", n_fields))
)

})

}

```

# Post-estimation analysis
```{r}
cf_results <- readRDS(here::here("Results/yield_response_function_for_one_iteration/YRF_CF_1fields.rds"))
```

## Visualization of yield response curve fit

```{r}
test_data <- cf_results$test_data
train_data <- cf_results$train_data
response_data <- cf_results$response_data

true_response <-
  test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
  .[, ] %>%
  expand_grid_df(., data.table(N = train_data[, seq(min(N), max(N), length = 100)])) %>%
  .[, yield := gen_yield_QP(b0, b1, b2, Nk, N)]

check_id <- test_data[order(Nk), aunit_id] %>%
  .[seq(1, nrow(test_data), length = 5)] %>%
  .[-1]

cf_curves <-
  response_data[aunit_id %in% check_id, ] %>% .[, type := "Estiamated by CF"]

true_curves <-
  true_response[aunit_id %in% check_id, ] %>%
  .[, type := "True"] %>%
  .[, .(aunit_id, N, yield, type)] %>%
  setnames("N", "T")

yield_intercept <- true_curves[, .(yield = mean(yield)), by = aunit_id]

yield_adjustment <-
  cf_curves[, .(est = mean(est)), by = aunit_id] %>%
  yield_intercept[., on = "aunit_id"] %>%
  .[, yield_adj := yield - est] %>%
  .[, .(aunit_id, yield_adj)]

cf_curves <-
  yield_adjustment[cf_curves, on = "aunit_id"] %>%
  .[, yield := est + yield_adj] %>%
  .[, .(aunit_id, yield, T, type)]

curves_data_il <-
  rbind(cf_curves, true_curves) %>%
  .[, type := factor(type, levels = c("True", "Estiamated by CF"))] %>%
  .[, group := .GRP, by = aunit_id]

g_response_estimated <-
  ggplot() +
  geom_line(
    data = curves_data_il[T <= 260, ],
    aes(
      y = yield,
      x = T,
      color = factor(group),
      linetype = type
    )
  ) +
  scale_color_discrete(name = "aunit_id Number") +
  scale_linetype_discrete(name = "Curve Type") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ylab("Corn Yield (bu/acre)") +
  xlab("Nitrogen Application Rate (lb/acre)")

# saveRDS(
#   object = g_response_estimated,
#   file   = here::here("Results", "g_response_estimated.rds")
# )

```

# All YRF

```{r}
#--- libraries ---#
library(pasim)
library(data.table)
library(dplyr)
library(ggplot2)
library(here)
library(purrr)
library(tidyr)     
library(patchwork)

process_cf_file <- function(file_path, label) {
  cf_results <- readRDS(file_path)

  test_data     <- cf_results$test_data
  train_data    <- cf_results$train_data
  response_data <- cf_results$response_data

  # true response curves
  true_response <- test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
    tidyr::expand_grid(N = seq(min(train_data$N),
                               max(train_data$N),
                               length = 100)) %>%
    as.data.table() %>%
    .[, yield := gen_yield_QP(b0, b1, b2, Nk, N)]

  # pick a few ids
  # check_id <- test_data[order(Nk), aunit_id] %>%
  #   .[seq(1, nrow(test_data), length = 2)] %>%
  #   .[-1]
   # Make CF use the same aunit_id as SO_ANN visualization
so_ann_csv <- here::here(
  "results", "yield_response_function_for_one_iteration",
  "YRF_SO_ANN_1_fields",   
  "yield_response_1.csv"
)

if (file.exists(so_ann_csv)) {
  so_df <- readr::read_csv(so_ann_csv, show_col_types = FALSE)
  # choose the same ID that SO_ANN used (the first unique aunit_id in its CSV)
  shared_id <- unique(so_df$aunit_id)[1]

  # ensure it exists in CF test_data
  if (shared_id %in% test_data$aunit_id) {
    check_id <- shared_id
  } else {
    # fallback: use the first available CF id
    check_id <- unique(test_data$aunit_id)[1]
  }
} else {
  # fallback if file missing
  check_id <- unique(test_data$aunit_id)[1]
}

message("Using aunit_id = ", check_id, " to match SO_ANN")

  # estimated by CF
  cf_curves <- response_data[aunit_id %in% check_id, ] %>%
    .[, type := "Estimated by CF"]

  # true curves
  true_curves <- true_response[aunit_id %in% check_id, ] %>%
    .[, type := "True"] %>%
    .[, .(aunit_id, N, yield, type)] %>%
    setnames("N", "T")

  # adjust intercepts
  yield_intercept <- true_curves[, .(yield = mean(yield)), by = aunit_id]
  yield_adjustment <- cf_curves[, .(est = mean(est)), by = aunit_id] %>%
    yield_intercept[., on = "aunit_id"] %>%
    .[, yield_adj := yield - est] %>%
    .[, .(aunit_id, yield_adj)]

  cf_curves <- yield_adjustment[cf_curves, on = "aunit_id"] %>%
    .[, yield := est + yield_adj] %>%
    .[, .(aunit_id, yield, T, type)]

  # combine
  curves_data_il <- rbind(cf_curves, true_curves) %>%
    .[, type := factor(type, levels = c("True", "Estimated by CF"))] %>%
    .[, group := .GRP, by = aunit_id]

  # plot

  ggplot() +
  geom_line(
    data = curves_data_il[T <= 260, ],
    aes(y = yield, x = T, color = factor(group), linetype = type)
  ) +
  scale_color_discrete(name = "aunit_id Number") +
  scale_linetype_discrete(name = "Curve Type") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ylab("Corn Yield (bu/acre)") +
  xlab("Nitrogen Application Rate (lb/acre)") +
  ggtitle(paste("CF -", label))

}

# file_labels <- c("1fields", "3fields", "5fields", "10fields", "20fields")
# file_paths <- here::here(
#   "results/yield_response_function_for_one_iteration",
#   paste0("YRF_CF_", file_labels, ".rds")
# )

field_counts <- c(1, 3, 5, 10, 20)

file_paths <- here::here(
  "results/yield_response_function_for_one_iteration",
  sprintf("YRF_CF_%dfields.rds", field_counts)
)

file_labels <- sprintf("%d fields", field_counts)

#---all plots ---#
all_plots <- map2(file_paths, file_labels, process_cf_file)

# saveRDS(
#   all_plots,
#   here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF_double.rds")
# )

```

# Visualize YRF

```{r}

all_plots <- readRDS(here::here("results/yield_response_function_for_one_iteration", "all_YRF_plots_CF_double.rds"))

# Example one plot
#all_plots[[1]]

# arrange 
wrap_plots(all_plots, ncol = 2)

```