### 2025-11-17 codes/R/GRF_CF_Loop.qmd

---
title: "GRF CF Loop"
format:
  html:
    number-sections: true
    number-depth: 1
    theme: flatly
    toc: true
execute:
  echo: true
  message: false
  warning: false
---

## Packages and functions
```{r setup_not_remove, echo=FALSE, include = F}
#--- packages ---#
library(tidyverse)
library(data.table)
library(mgcv)
library(spatialsample)
library(parallel)
library(sf)
library(grf)
library(ranger)
library(xgboost)
library(mgcv)
library(here)
library(reticulate)
library(readr)
library(dplyr)
library(tidyr)

Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  VECLIB_MAXIMUM_THREADS = "1" 
)

#--- R functions ---#
source(here("codes", "R", "Utility.R"))

```

## load data
```{r}
#raw_sim_data <- readRDS(here::here("data/raw_sim_data.rds"))

# how many fields to train on?
n_fields <- 5

splits_file <- here(
  "data", "train_test_split",
  sprintf("train_test_splits_%dfields.csv", n_fields)
)
output_dir <- here(
  "results/GRF_CF",
  sprintf("CF_outcome_%d_fields", n_fields)
)
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

progress_csv <- file.path(output_dir, "progress.csv")

# ─── LOAD SPLITS ────────────────────────────────────────────────────────────────
train_test_splits <- read_csv(splits_file)

```

# CF loop
```{r}
# ── EONR subfolder ────────────────────────────────────────────────
eonr_dir <- file.path(output_dir, "eonr")
if (!dir.exists(eonr_dir)) dir.create(eonr_dir, recursive = TRUE)

## ── Loop over test sims ───────────────────────────────────────────────────────
#for (i in seq_len(nrow(train_test_splits)))
#for (i in seq_len(1))
for (i in seq_len(nrow(train_test_splits))) {
  test_sim_id <- train_test_splits$test_id[i]
  out_fn <- file.path(output_dir, sprintf("sim_%03d.rds", as.integer(test_sim_id)))

  # skip if already saved
  if (file.exists(out_fn)) {
    message("Skipping sim ", test_sim_id, " (already saved)")
    next
  }

  tryCatch(
    {
      ## 1) train/test IDs
      split_row     <- train_test_splits[i, ]
      train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:n_fields)]))
      sim_id_ls     <- c(train_sim_ids, test_sim_id)

      ## 2) train & test data
      data <- raw_sim_data$reg_data[[1]] %>%
        .[sim %in% sim_id_ls, ] %>%
        .[, .(sim, data)] %>%
        tidyr::unnest(cols = c(data)) %>%
        as.data.table()

      train_data <- data[sim != test_sim_id, ]
      test_data  <- raw_sim_data$reg_data[[1]] %>%
        .[sim == test_sim_id, ] %>%
        .[, .(sim, data)] %>%
        tidyr::unnest(cols = c(data)) %>%
        as.data.table()

      ## 3) CF 
      x_vars <- c("Nk", "plateau", "b0")
      T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
      Y      <- train_data[, yield]
      X      <- as.matrix(train_data[, ..x_vars])
      W      <- X
      X_test <- as.matrix(test_data[, ..x_vars])

      ## 4) run cf in python
      Sys.setenv(
        OMP_NUM_THREADS      = "1",
        MKL_NUM_THREADS      = "1",
        KMP_AFFINITY         = "disabled",
        KMP_DUPLICATE_LIB_OK = "TRUE"
      )

      te_hat_cf <- run_CF_c_py(
        Y,
        T_info$T_sp,
        X,
        W,
        n_estimators = 2000
      )
      te_info <- get_te(te_hat_cf, test_data, x_vars, "aunit_id")

      ## 5) Response curves
      T_seq <- train_data[, seq(min(N), max(N), length = 200)]
      response_data <- find_response_semi(T_seq, T_info, te_info)

      ## 6) Site-specific EONR
      pCorn <- 6.25 / 25.4
      pN    <- 1 / 0.453592

      ss_eonr <- response_data %>%
        .[, profit := est * pCorn - pN * T] %>%
        .[, .SD[which.max(profit), ], by = aunit_id] %>%
        .[, .(aunit_id, opt_N_hat = T)]

      true_ss_eonr <- test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
        .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
        .[, opt_N := pmin(Nk, opt_N)] %>%
        .[, opt_N := pmax(0,  opt_N)] %>%
        .[, .(aunit_id, opt_N)]

      combined_eonr <- ss_eonr[true_ss_eonr, on = "aunit_id"]
      corr_val <- combined_eonr[, cor(opt_N, opt_N_hat)]

      ## 7) SAVE EONR 
      e <- as.data.table(combined_eonr)[, .(aunit_id, opt_N_hat, opt_N)]
      e[, sim := as.integer(test_sim_id)]
      data.table::setcolorder(e, c("aunit_id", "opt_N_hat", "opt_N", "sim"))

      fn_eonr   <- file.path(eonr_dir, sprintf("sim_%03d_eonr.rds", as.integer(test_sim_id)))
      tmp_eonr  <- paste0(fn_eonr, ".tmp")
      saveRDS(e, tmp_eonr, compress = "xz")
      file.rename(tmp_eonr, fn_eonr)

      # optional
      csv_path <- file.path(eonr_dir, "eonr_progress.csv")
      data.table::fwrite(e, csv_path, append = file.exists(csv_path))

      ## 8) save the full result object 
      # res <- list(
      #   te_hat_cf     = te_hat_cf,
      #   te_info       = te_info,
      #   T_seq         = T_seq,
      #   response_data = response_data,
      #   test_data     = test_data,
      #   train_data    = train_data,
      #   combined_eonr = combined_eonr,
      #   corr          = corr_val
      # )
      # 
      # tmp_fn <- paste0(out_fn, ".tmp")
      # saveRDS(res, tmp_fn, compress = "xz")
      # file.rename(tmp_fn, out_fn)

      ## 9) Log progress
      prog <- data.frame(
        test_sim_id = test_sim_id,
        corr        = corr_val,
        n_train     = nrow(train_data),
        saved_file  = out_fn,
        eonr_file   = fn_eonr,
        ts          = as.character(Sys.time())
      )
      write.table(
        prog,
        file      = progress_csv,
        sep       = ",",
        row.names = FALSE,
        col.names = !file.exists(progress_csv),
        append    = TRUE
      )

      message(
        "Finished sim ", test_sim_id, " → ", basename(out_fn),
        " | eonr: ", basename(fn_eonr),
        " (corr = ", round(corr_val, 3), ")"
      )

      # Free ram 
      rm(te_hat_cf, te_info, response_data, train_data, test_data, e)
      if (i %% 10 == 0) gc()
    },
    error = function(e) {
      message("ERROR in sim ", test_sim_id, ": ", conditionMessage(e))
      err_row <- data.frame(
        test_sim_id = test_sim_id,
        error       = conditionMessage(e),
        ts          = as.character(Sys.time())
      )
      write.table(
        err_row,
        file      = file.path(output_dir, "errors.csv"),
        sep       = ",",
        row.names = FALSE,
        col.names = !file.exists(file.path(output_dir, "errors.csv")),
        append    = TRUE
      )
    }
  )
}

```

```{r}

# ── EONR subfolder ──────────────────────────────────────────────────────────────
eonr_dir <- file.path(output_dir, "eonr")
if (!dir.exists(eonr_dir)) dir.create(eonr_dir, recursive = TRUE)

## ── Loop over test sims ───────────────────────────────────────────────────────
#for (i in seq_len(1))
#for (i in seq_len(nrow(train_test_splits)))
for (i in seq_len(1)) {
  test_sim_id <- train_test_splits$test_id[i]
  out_fn <- file.path(output_dir, sprintf("sim_%03d.rds", as.integer(test_sim_id)))

  if (file.exists(out_fn)) {
    message("Skipping sim ", test_sim_id, " (already saved)")
    next
  }

  tryCatch({
    ## 1) train/test IDs
    split_row     <- train_test_splits[i, ]
    train_sim_ids <- as.integer(unlist(split_row[paste0("train_", 1:n_fields)]))
    sim_id_ls     <- c(train_sim_ids, test_sim_id)

    ## 2) train & test data
    data <- raw_sim_data$reg_data[[1]] %>%
      .[sim %in% sim_id_ls, ] %>%
      .[, .(sim, data)] %>%
      tidyr::unnest(cols = c(data)) %>%
      as.data.table()

    train_data <- data[sim != test_sim_id, ]
    test_data  <- raw_sim_data$reg_data[[1]] %>%
      .[sim == test_sim_id, ] %>%
      .[, .(sim, data)] %>%
      tidyr::unnest(cols = c(data)) %>%
      as.data.table()

    ## 3) CF variables
    x_vars <- c("Nk", "plateau", "b0")
    T_info <- prepare_T_mat(formula(yield ~ s(N, k = 4), m = 2), data = train_data)
    Y      <- train_data[, yield]
    X      <- as.matrix(train_data[, ..x_vars])
    W      <- X
    X_test <- as.matrix(test_data[, ..x_vars])

    ## 4) cf in GRF
    cf_fit <- causal_forest(
      X = X,
      Y = Y,
      W = T_info$T_sp,     # spline
      num.trees = 2000,
      min.node.size = 10,
      seed = 3747823
    )

    # Estimate treatment effects for test data
    te_hat_cf <- predict(cf_fit, newdata = X_test)$predictions

    te_info <- data.table(aunit_id = test_data$aunit_id, est = te_hat_cf)

    ## 5) Response curves
    T_seq <- train_data[, seq(min(N), max(N), length = 200)]
    response_data <- find_response_semi(T_seq, T_info, te_info)

    ## 6) Site-specific EONR
    pCorn <- 6.25 / 25.4
    pN    <- 1 / 0.453592

    ss_eonr <- response_data %>%
      .[, profit := est * pCorn - pN * T] %>%
      .[, .SD[which.max(profit), ], by = aunit_id] %>%
      .[, .(aunit_id, opt_N_hat = T)]

    true_ss_eonr <- test_data[, .(aunit_id, b0, b1, b2, Nk)] %>%
      .[, opt_N := (pN / pCorn - b1) / (2 * b2)] %>%
      .[, opt_N := pmin(Nk, opt_N)] %>%
      .[, opt_N := pmax(0,  opt_N)] %>%
      .[, .(aunit_id, opt_N)]

    combined_eonr <- ss_eonr[true_ss_eonr, on = "aunit_id"]
    corr_val <- combined_eonr[, cor(opt_N, opt_N_hat)]

    ## 7) SAVE EONR
    e <- as.data.table(combined_eonr)[, .(aunit_id, opt_N_hat, opt_N)]
    e[, sim := as.integer(test_sim_id)]
    data.table::setcolorder(e, c("aunit_id", "opt_N_hat", "opt_N", "sim"))

    fn_eonr   <- file.path(eonr_dir, sprintf("sim_%03d_eonr.rds", as.integer(test_sim_id)))
    tmp_eonr  <- paste0(fn_eonr, ".tmp")
    saveRDS(e, tmp_eonr, compress = "xz")
    file.rename(tmp_eonr, fn_eonr)

    csv_path <- file.path(eonr_dir, "eonr_progress.csv")
    data.table::fwrite(e, csv_path, append = file.exists(csv_path))

    ## 8) Save results object (optional)
    # res <- list(
    #   cf_fit        = cf_fit,
    #   te_info       = te_info,
    #   T_seq         = T_seq,
    #   response_data = response_data,
    #   test_data     = test_data,
    #   train_data    = train_data,
    #   combined_eonr = combined_eonr,
    #   corr          = corr_val
    # )
    # tmp_fn <- paste0(out_fn, ".tmp")
    # saveRDS(res, tmp_fn, compress = "xz")
    # file.rename(tmp_fn, out_fn)

    ## 9) Log progress
    prog <- data.frame(
      test_sim_id = test_sim_id,
      corr        = corr_val,
      n_train     = nrow(train_data),
      saved_file  = out_fn,
      eonr_file   = fn_eonr,
      ts          = as.character(Sys.time())
    )
    write.table(
      prog,
      file      = progress_csv,
      sep       = ",",
      row.names = FALSE,
      col.names = !file.exists(progress_csv),
      append    = TRUE
    )

    message("Finished sim ", test_sim_id,
            " | corr = ", round(corr_val, 3))

    # Free RAM
    rm(cf_fit, te_info, response_data, train_data, test_data, e)
    if (i %% 10 == 0) gc()
  },
  error = function(e) {
    message("ERROR in sim ", test_sim_id, ": ", conditionMessage(e))
    err_row <- data.frame(
      test_sim_id = test_sim_id,
      error       = conditionMessage(e),
      ts          = as.character(Sys.time())
    )
    write.table(
      err_row,
      file      = file.path(output_dir, "errors.csv"),
      sep       = ",",
      row.names = FALSE,
      col.names = !file.exists(file.path(output_dir, "errors.csv")),
      append    = TRUE
    )
  })
}

```