### 2025-11-25 presentation/proposal/proposal.qmd
Edit proposal

---
title: "Proposal Defense"
author:
  - name: Mona Mousavi
institute:
  - University of Nebraska Lincoln
date: 06/12/2025
format:
  revealjs:
    theme: [default, ./../custom.scss]
    fontsize: 1.4em
    callout-icon: false
    scrollable: true
    fig-dpi: 400
  beamer:
    theme: Singapore
    navigation: horizontal
    section-titles: false
    slide-level: 2
    fontsize: "10pt"
logo: nebraska-n.jpg
---

```{r, include = FALSE}
library(terra)
library(sf)
library(tidyterra)
library(ggplot2)
library(data.table)
library(dplyr)
library(flextable)
```

# Chapter 1: A new model selection approach based on local economically optimal input rate

## Background

+ Uniform input management strategies can lead to both economic losses and environmental inefficiencies.

+ Variable-rate input management as a solution (accounts for spatial heterogeneity in crop input demand).

+ On-farm precision experiments (OFPE) as a promising way (GPS technologies and variable-rate applicators).
  + Run on-farm agronomic trials and collect data
  + <span style = "color: blue;">Estimate how yield responds to nitrogen based on observed characteristics</span>
  + Identify site-specific optimal nitrogen rates

## Motivation

+ Machine learning approaches being increasingly used for estimating site-specific yield response functions
  + identify the best model among several candidate ML approaches and also tuning hyper parameters via cross-validation as the key part of such process.

+ In all of the previous studies, the best model was selected based on their abilities to predict <span style = "color: blue;">yield level</span> at the observation points (yield-based model selection).

+ However, the ultimate goal is not to predict yield levels, but to estimate yield response (causal marginal impact of nitrogen on crop yield).

## Research Objective

+ Proposes an alternative to conventional yield-based model selection. Specifically, we select models based on their ability to predict local EONR using spatial cross-validation.

+ Evaluate this approach, through Monte Carlo simulations, and compare its performance against the standard yield-based selection method.

## Method: Overview

Monte Carlo simulations with 500 iterations.

### Steps for a single iteration:

  + Step 1: Generate OFPE experiment data
  + Step 2: Identify the true best (most profitable VRA application rate) model among the five candidate models
  + Step 3: Select the best model among the 5 candidate models based on their abilities to predict
    + yield level (conventional approach)
    + local-EONR (our approach)
  + Step 4: Evaluate the model selection approaches

### The five candidate models:

+ Linear model 
+ Linear spatial error (SE) model 
+ Random forest (RF) 
+ Boosted regression forest (BRF) 
+ Causal forest (CF)

## Method: Data generation

+ The data generated from a simulated corn production field (about 37.3 hectares) with spatially varying field characteristics.

+ One possible scenario of the spatial variability in the field characteristics is shown below.

+ This study assumed the yield data were collected at the subplot level (total 1440 subplots).

```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "95%"
knitr::include_graphics(
  here::here("presentation/proposal/g_field_map.png")
)
```

## Method: True yield responses

The true yield is assumed to follow the quadratic plateau functional form as follows:

$$
y_{j, i}=f_{j, i}(N)= \begin{cases}\alpha_{j, i}+\beta_{j, i} N+\gamma_{j, i} N^2+\varepsilon_{j, i}, & N<\tau_{j, i} \\ \eta_{j,i}+\varepsilon_{j, i}, & N \geq \tau_{j, i}\end{cases}
$$

+ $j \in{1,\ 2,\ 3,...,1440}$: the subplot,
+ $i \in{1,\ ...,6}$: the cell within a subplot,
+ $y_{j,i}$: Corn yield,
+ $N$: Nitrogen rate,
+ $\alpha, \beta, \gamma$: parameters (soil/field characteristics) that vary within a field in a spatially correlated manner and govern how yield responds to $N$
+ $\tau_{j, i}$:  the $N$ rate at which the yield reaches a plateau.
+ $\eta_{j,i}$: the yield at $N = \tau_{j, i}$,
+ $\varepsilon_{j, i}$: spatial autocorrelation.

## Method: On-farm trial (nitrogen rate)

Six N trial rates ($\{80, 130, 154, 184, 219, 270\}$ kg/ha.) following a 6 × 6 Latin square design.

(Note: 8 blocks, each block including 36 plots ($8*36=288$ plots) to establish the Latin square design).

```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "100%"
knitr::include_graphics(
  here::here("presentation/proposal/g_N.png")
)
```

## Method: Yield data

Based on the response parameters $\left(\alpha_{j, i}, \beta_{j, i}, \gamma_{j, i}, \tau_{j, i}, \eta_{j,i}\right)$ and error term $\varepsilon_{j,i}$, the yield $y_{j, i}$ for each cell $(j, i)$ was generated following the true response function.

* Note:
The harvesting equipment was assumed to be able to accurately collect yield data only at the subplot resolution. Therefore, the observed yield data were at the resolution of subplot level.

```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "100%"
knitr::include_graphics(
here::here("presentation/proposal/g_y.png")
)
```

## Method: Independent variables

+ The true response parameters $\left(\alpha_{j, i}, \beta_{j, i}, \gamma_{j, i}, \tau_{j, i}, \eta_{j,i}\right)$ are not directly observable by producers or researchers.

+ This study creates some covariates (correlated with those true parameters) to serve as the observable field characteristics:

$$X_j = \left(\beta^{1,1}_j, \beta^{1,2}_j, \beta^{2,1}_j, \beta^{2,2}_j, \tau^{1,1}_j, \tau^{1,2}_j, \tau^{2,1}_j, \tau^{2,2}_j, \eta^{1,1}_j, \eta^{1,2}_j, \eta^{2,1}_j, \eta^{2,2}_j \right)$$

$$\beta = \beta^{1,1} + \beta^{1,2} + \beta^{2,1} + \beta^{2,2}$$ 
$$\tau = \tau^{1,1} + \tau^{1,2} + \tau^{2,1} + \tau^{2,2}$$ 
$$
\eta = \min(\min(\eta^{1,1}, \eta^{1,2}), \min(\eta^{2,1}, \eta^{2,2}))
$$

## Identifying the true best model among the five candidate models

+ Each of the candidate models are trained using the whole experiment data and then variable rate EONR are estimated for the whole field.

+ The estimated variable rate N recommendations are evaluated against the true variable rate N to evaluate the performance of the models (using profit loss relative to the true site-specific EONR as the criteria):

The profit loss of model $m$:

$$
\Delta\pi_m = \pi_{true} - \pi_m
$$

Where:

$$
\pi_{true} = \frac{1}{N} \cdot \sum_{i=1}^{N} \left[P_c \cdot f(N^*_i, X_i) - P_N \cdot N^*_i\right]
$$

$$
\pi_{m^*} = \frac{1}{N} \cdot \sum_{i=1}^{N} \left[P_c \cdot f(\widehat{N}^*_{m}(X_i), X_i) - P_N \cdot \widehat{N}^*_{m}(X_i)\right]
$$

## Model Selection

+ For both approaches (yield prediction accuracy, local EONR prediction accuracy) spatial cross-validation is employed.
+ The entire dataset is randomly split into a train and test dataset in a spatially clustered manner, which is repeated to create multiple folds.

```{r}
knitr::include_graphics("g_split_example.png")
```

## Model selection based on yield level prediction

For each fold:

1. train the model using the train data of the fold
2. predict yield for the test data based on the trained model ($\hat{y}_{i,f}$)
3. calculate RMSE of yield prediction as 
$$
RMSE\text{-}Yield_{m,f} = \sqrt{\frac{\sum_{i=1}^{N_t}(\hat{y}_{i,f} - y_{i,f})^2}{N_t}}
$$

## Model selection based on local EONR

+ True EONRs are never observable in the real world

+ Using estimated “local uniform EONR” as the proxy for the true local EONR  to overcome this challenge

+ Steps for each fold and candidate models:
  1. train the GAM model using the **test** data 
  2. estimate uniform EONR for the **test** data ($\hat{N}^*_{f,gam}$)
  3. train the candidate model using the **train** data 
  4. estimate site-specific EONR for the **test** data based on the trained candidate model ($\hat{N}^*_{i, f,m}$). 
  5. Calculate the average of the site-specific EONR ($\hat{N}^*_{f,m}$)
  6. Calculate the RMSE of the estimated local EONR:

$$
RMSE\text{-}LE_{m} = \sqrt{\frac{\sum_{f=1}^{F} (\hat{N}^*_{f,m} - \hat{N}^*_{f,gam})^2}{F}}
$$

## Evaluation of model selection performance

+ Selected models (based on both approaches) are evaluated based on the following difference:

$$
\pi_{m^*} - \pi_{m^s}
$$

+ $\pi_{m^*}$: the highest profit that could be achieved if (EONR from) the true best model were used. 
+ $\pi_{m^s}$: the profit achieved by the selected model

## Simulation Results: Performance of the model selection approaches

```{r}
#| fig-align: "center"
#| out-width: "90%"
knitr::include_graphics("g_ranking.png")
```

## Simulation Results: Accuracy of the model selection approaches

```{r}
#| echo: false
#| results: 'asis'
tbl <- readRDS("main_selection_accuracy_table.rds")
invisible(flextable::save_as_image(tbl, path = "table_accuracy.png"))
knitr::include_graphics("table_accuracy.png")

```

## Simulation Results: profit loss associated with the seleted models relative to the actual best models

```{r}
#| fig-align: "center"
#| out-width: "90%"
knitr::include_graphics("g_pi_loss.png")
```

## Simulation Results: Accuracy of Local EONR estimation with GAM

```{r}
#| fig-align: "center"
#| out-width: "90%"
knitr::include_graphics("g_gam_leonr.png")
```

## Conclusion

This study demonstrates the economic advantage of local EONR method in making more accurate and cost-effective input management decisions compared to the yield-based model selection.

# Chapter 2: Causal methods for Site-specific yield response function estimation

## Background

+ Efficient nitrogen management remains one of the most critical challenges in modern agricultural production.

+ Applying right amount of nitrogen at right place (and right time) recognizing the need for nitrogen based on observed characteristics (soil N, soil and field characteristics, weather).

+ An increasing amount of data from on-farm field trials has been accumulated over the years.

+ Estimating yield response functions using data pooled from multiple fields can provide significant agronomic and economic insights.

## Motivation

+ When on-farm trial data is pooled, the treatment variable becomes "continuous" rather than discrete.

+ Existing approaches—such as direct applications of random forests or neural networks—often tend to fail to capture heterogeneous treatment effects of a continuous variable.

+ Most causal machine learning methods (e.g., R-learner) are designed for binary treatments, whereas our objective is to recover the full yield response <span style = "color: blue;">curve</span>.

## Research Objective

+ Find a causal machine learning approach capable of effectively identifying heterogeneous treatment effects tailored for yield response function estimation using data from on-farm trials, where the treatment variable is continuous and randomized.

+ Evaluate the performance of the proposed method through Monte Carlo simulations, comparing it against commonly used existing approaches.

## Method: Double-ML (binary treatment ($T$) case):

DML works fine in a <span style = "color: blue;">binary</span> treatment ($T$) case (model assumes linearity in $T$):

```{=tex}
\begin{align*}
\mbox{Model 1: } Y = \theta(X)\cdot T + g(X, W) + v
\end{align*}
```

+ $Y$: Dependent variable 
+ $T$: Binary treatment variable (variable of interest) 
+ $X$: Set of variables that affect the impact of $T$ and also $Y$ directly
+ $\theta(X)$: Treatment effect as a function of $X$
+ $W$: Set of variables that affect $Y$ directly
+ $v$: Error term

---

$$
\begin{aligned} & Y_{res}=Y-E[Y \mid X, W]=Y-f(X, W) \\ & T_{res}=T-E[T \mid X, W]=T-h(X, W)\end{aligned}
$$

+ A simple regression of $Y_{res}$ on $T_{res}$ can recover $\theta(X)$:

$$
\begin{aligned}
& Y_{res} = \\
& \begin{aligned}
& {[\theta(X)\cdot T + g(X,W) + v] - [E\{ \theta(X)\cdot T + g(X,W) + v \mid X,W\}]= } \\
& {[\theta(X)\cdot T + g(X,W) + v] - [\theta(X)\cdot E(T\mid X,W) + g(X,W)]=} \\
&  {[\theta(X)\cdot T + g(X,W) + v] - [\theta(X)\cdot h(X,W) + g(X,W)]=} \\
&  \theta(X)\cdot [T - h(X,W)] + v =  \\
& \theta(X)\cdot T_{res} + v
\end{aligned}
\\[0.5em]
& \Rightarrow\quad Y_{res} = \theta(X)\cdot T_{res} + v
\end{aligned}
$$

## Method: Double-ML (continuous treatment ($T$) case):

DML fails in a <span style = "color: blue;">continuous</span> treatment ($T$) case:

```{=tex}
\begin{align*}
\mbox{Model 2: } Y = \theta(X, T) + g(X, W) + v
\end{align*}
```

+ $T$: <span style = "color: blue;">Continuous </span> treatment variable (e.g., nitrogen rate) 
+ $\theta(X, T)$: Impact of $X$ and $T$ with their full interactions allowed

$\begin{aligned}
& Y_{res}= \\
& \theta(X, T)+g(X, W)+v-E\{\theta(X, T)+g(X, W)+v \mid X, W\}= \\
& \theta(X, T)+g(X, W)+v-E(\theta(X, T) \mid X, W)-g(X, W) \\
& \Rightarrow Y_{res}=\theta(X, T)-E(\theta(X, T) \mid X, W)+v
\end{aligned}$

So, orthogonalizing $T$ does not work: $\widehat{\theta}(X, T - g(X, W))$ cannot be used to approximate $E[\theta(X, T)]$.

## Our proposed approaches

+ Single-orthogonalization
+ Extended CF

## Single-orthogonalization

```{=tex}
\begin{align*}
\mbox{Model 2: } Y = \theta(X, T) + g(X, W) + v
\end{align*}
```

1. Regress $Y$ on $X$ and $W$ and predict $E[Y|X,W]$ (denoted as $f(X,W)$)
2. Regress $Y-f(X,W)$ on $T$

This approach works because The orthogonalization shifts the attention from the variations in $Y$ to the variations in $Y-f(X,W)$ just due to $X$ and $T$.

## Extended CF

+ We propose the use of smoothing splines. In spline regression, a single variable can be decomposed into multiple basis functions in a linear-in-parameter manner.

```{=latex}
\begin{align} \label{eq-spline-cf}
y = \sum_{k=1}^K \theta_k(X)\cdot S_k(T)+ g(X, W) + v
\end{align}
```

+ $S_k(T)$: the value of the basis function $S_k(\cdot)$ evaluated at $T$.

## Models

$$
y = f(X, N) + g(X,W) + v
$$

+ $Y$: Crop yield
+ $N$: Nitrogen rate (lb/acre) 
+ $X$: Collection of variables ($\alpha$, $\beta$, $\gamma$, and $\tau$)
+ $v$: Error term

## Scenarios

+ Estimation approaches:

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("f_est.png")
```

+ For each of 500 simulation rounds, one field is selected as the test field. The remaining fields are combined to construct the training set in five scenarios: 
  + number of fields from which data is combined: 1, 3, 5, 10 and 20

## Site-specific EONR estimation

For a given price of corn and nitrogen EONR is found by solving the following profit maximization problem:

$$
\widehat{\mathrm{EONR}}=\arg \max _N\left[P_{\text {Corn }} \times \hat{f}(X, N)-P_N \times N\right]
$$

+ $\widehat{f}$: yield response curve estimated by the model under evaluation

## Model evaluation

Model evaluation is based on the accuracy of predicting <span style = "color: blue;"> site-specific economically optimal nitrogen rates </span>($\widehat{EONR}$).

$$
RMSE_{EONR} = \sqrt{\frac{\sum_{i=1}^{n} (\widehat{EONR}_{i} - {EONR}_{i})^2}{n}}
$$

# Results and Conclusions

## Simulation (preliminary) results

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("fig_rmse_hist.png")
```

## Simulation (preliminary) results

```{r}
#| fig-align: "center"
#| out-width: "45%"
knitr::include_graphics("tbl_best_model_by_field_count.png")
```

## Simulation (preliminary) results

```{r}
#| echo: false
#| fig-width: 4
#| out-width: 4in
knitr::include_graphics("model_improvement.png")
```

## Conclusion

By combining orthogonalization with flexible final-stage learners, we demonstrate how to preserve the strengths of causal inference while accommodating continuous treatments and agronomic heterogeneity.

# Chapter 3: Economic assessment of biochar as a means to enhance soil water holding capacity

## Background

+ Improving soil health is considered to be an important means of enhancing crop production resilience to climate change
+ Soil water holding capacity (WHC) is an important property of soil. Higher soil WHC leads to
  + withstanding droughts better
  + less irrigation (conserving water resources)
+ Biochar is actively researched as a supplement to enhance soil WHC of production fields, but is expensive

## Research question and objective

:::{.callout-note title="Research Question"}
+ Is biochar economically viable?
+ If not, how much cost reduction in biochar needs to be achieved to make it economically viable?
:::

## Method

1. Simulate corn production using APSIM (a crop growth model) at different values of soil WHC for
   1. historical weather conditions 
   2. weather scenarios under climate change
2. Calculate profit
3. Identify the value of soil WHC enhancement

## Expected Outcomes